### 2015-03-24 OECD-TiVa-LMIC-GVCs.Rmd
layout

---
title: "OECD TiVa LMIC GVCs"
author: "Victor Kummritz and Bastiaan Quast"
date: "March 24, 2015"
output: ioslides_presentation
---

## Motivation

 - Global Value Chains relevance
 - GVCs in developement (capacity building)
 - General GVC picture overemphasises north-north because of higher monetary value

## This paper

## Data

ICIOs from TiVa (not WIOD)

TiVa data
 - 57 countries
 - 34 sectors
 - LMICs

## New Toolbox and Indicators

 - source / using countries specific indicators
 - decompor and gvc functions

## General Stylised Facts

 - section 3.1 of wwz
 - growth of GVC trade
 - length of GVC
 - determinants of an industry's participation
 - determinants of a country's participation
 - trade balanses in VA terms differ substantially from those in gross terms

## GVCs of LMICs

 - stylised facts for LMICs
 - source / using countries specific indicators for LMICs
 - nrca and trca of LMIC
 - fvafin vs. fvaint for LMICs (WWZ 3.3)
 - dvix decomposition (WWZ 3.4)

## Conclusion

### 2015-03-27 R/decomposition.R
add files for nrca

@@ -1,0 +1,66 @@
+# decomposition.R
+# Bastiaan Quast
+# bquast@gmail.com
+
+# load the library
+library(decompr)
+
+# run the Leontief decomposition
+l1995 <- decomp(wid1995,
+                wfd1995,
+                countries,
+                sectors,
+                output1995,
+                method = "leontief")
+l2000 <- decomp(wid2000,
+                wfd2000,
+                countries,
+                sectors,
+                output2000,
+                method = "leontief")
+l2005 <- decomp(wid2005,
+                wfd2005,
+                countries,
+                sectors,
+                output2005,
+                method = "leontief")
+l2008 <- decomp(wid2008,
+                wfd2008,
+                countries,
+                sectors,
+                output2008,
+                method = "leontief")
+
+# save Leontief the data
+save(l1995, l2000, l2005, l2008, file = "data/TiVa-Leontief.RData")
+
+
+
+# run the Wang-Wei-Zhu decomposition 
+w1995 <- decomp(wid1995,
+                wfd1995,
+                countries,
+                sectors,
+                output1995,
+                method = "wwz")
+w2000 <- decomp(wid2000,
+                wfd2000,
+                countries,
+                sectors,
+                output2000,
+                method = "wwz")
+w2005 <- decomp(wid2005,
+                wfd2005,
+                countries,
+                sectors,
+                output2005,
+                method = "wwz")
+w2008 <- decomp(wid2008,
+                wfd2008,
+                countries,
+                sectors,
+                output2008,
+                method = "wwz")
+
+# save the Wang-Wei-Zhu data
+save(w1995, w2000, w2005, w2008, file = "data/TiVa-WWZ.RData")

### 2015-03-27 R/import.R

@@ -1,0 +1,89 @@
+# import.R
+# Bastiaan Quast
+# bquast@gmail.com
+
+# load the readxl library
+# for excell files (.xls / .xlsx ) 
+library(readxl)
+
+# read the final demand sheets
+wfd1995 <- read_excel("data/WFD1.xlsx", sheet = 1, col_names = FALSE, skip = 2)
+wfd2000 <- read_excel("data/WFD1.xlsx", sheet = 2, col_names = FALSE, skip = 2)
+wfd2005 <- read_excel("data/WFD1.xlsx", sheet = 3, col_names = FALSE, skip = 2)
+wfd2008 <- read_excel("data/WFD1.xlsx", sheet = 4, col_names = FALSE, skip = 2)
+
+# separate out country and sector names
+countries <- wfd1995[ , 1]
+sectors   <- wfd1995[ , 2]
+
+# filter out doubles
+countries <- unique(countries)
+sectors   <- unique(sectors)
+
+# verify created objects
+print(countries)
+print(sectors)
+
+# remove invalid parts by subsetting only valid parts
+countries <- countries[1:58]
+sectors   <- sectors[1:34]
+
+# verify again
+print(countries)
+print(sectors)
+
+# create objects for number of countries and number of sectors
+# and multiple of the two
+G  <- length(countries)
+N  <- length(sectors)
+GN <- G * N
+
+# remove all data excel final demand statistics
+wfd1995 <- wfd1995[ 1:GN, -c(1:2) ]
+wfd2000 <- wfd2000[ 1:GN, -c(1:2) ]
+wfd2005 <- wfd2005[ 1:GN, -c(1:2) ]
+wfd2008 <- wfd2008[ 1:GN, -c(1:2) ]
+
+# save output separately
+output1995 <- wfd1995[ , length(wfd1995) ]
+output2000 <- wfd2000[ , length(wfd2000) ]
+output2005 <- wfd2005[ , length(wfd2005) ]
+output2008 <- wfd2008[ , length(wfd2008) ]
+
+# remove last column
+wfd1995 <- wfd1995[ , -length(wfd1995) ]
+wfd2000 <- wfd2000[ , -length(wfd2000) ]
+wfd2005 <- wfd2005[ , -length(wfd2005) ]
+wfd2008 <- wfd2008[ , -length(wfd2008) ]
+
+# read the ICIOs
+wid1995 <- read.table("~/OECD-ICIO-Export-Decomposition/data/Wicio34_1995_test.csv",
[…]

### 2015-03-27 R/leontief-matrix.R

@@ -1,0 +1,43 @@
+# leontief matrix form
+# Bastiaan Quast
+# bquast@gmail.com
+
+library(decompr)
+
+lm1995 <- load_tables_vectors(wid1995,
+                              wfd1995,
+                              countries,
+                              sectors,
+                              output1995)
+lm2000 <- load_tables_vectors(wid2000,
+                              wfd2000,
+                              countries,
+                              sectors,
+                              output2000)
+lm2005 <- load_tables_vectors(wid2005,
+                              wfd2005,
+                              countries,
+                              sectors,
+                              output2005)
+lm2008 <- load_tables_vectors(wid2008,
+                              wfd2008,
+                              countries,
+                              sectors,
+                              output2008)
+
+lmf1995 <- leontief(lm1995, long=FALSE)
+lmf2000 <- leontief(lm2000, long=FALSE)
+lmf2005 <- leontief(lm2005, long=FALSE)
+lmf2008 <- leontief(lm2008, long=FALSE)
+
+# run nrca's
+
+library(gvc)
+
+nrca1995 <- nrca(lmf1995)
+nrca2000 <- nrca(lmf2000)
+nrca2005 <- nrca(lmf2005)
+nrca2008 <- nrca(lmf2008)
+
+# save nrca's
+save(nrca1995, nrca2000, nrca2005, nrca2008, file ="data/nrca.RData")

### 2015-04-12 R/import.R
impute missing MLT c23 value

@@ -57,26 +57,29 @@ wfd2005 <- wfd2005[ , -length(wfd2005) ]
 wfd2008 <- wfd2008[ , -length(wfd2008) ]
 
 # read the ICIOs
-wid1995 <- read.table("~/OECD-ICIO-Export-Decomposition/data/Wicio34_1995_test.csv",
+wid1995 <- read.table("data/Wicio34_1995_test.csv",
                       sep=";",
                       quote="\"",
                       stringsAsFactors=FALSE,
                       skip = 2 )
-wid2000 <- read.table("~/OECD-ICIO-Export-Decomposition/data/Wicio34_2000_test.csv",
+wid2000 <- read.table("data/Wicio34_2000_test.csv",
                       sep=";",
                       quote="\"",
                       stringsAsFactors=FALSE,
                       skip = 2 )
-wid2005 <- read.table("~/OECD-ICIO-Export-Decomposition/data/Wicio34_2005_test.csv",
+wid2005 <- read.table("data/Wicio34_2005_test.csv",
                       sep=";",
                       quote="\"",
                       stringsAsFactors=FALSE,
                       skip = 2 )
-wid2008 <- read.table("~/OECD-ICIO-Export-Decomposition/data/Wicio34_2008_test.csv",
+wid2008 <- read.table("data/Wicio34_2008_test.csv",
                       sep=";",
                       quote="\"",
                       stringsAsFactors=FALSE,
                       skip = 2 )
+
+# find output MLT C23
+mltc23 <- wid1995[1982,1879]
 
 # remove first two rows with countries and sectors
 # and extra information at the bottom
@@ -84,6 +87,10 @@ wid1995 <- wid1995[ 1:GN , -c(1:2) ]
 wid2000 <- wid2000[ 1:GN , -c(1:2) ]
 wid2005 <- wid2005[ 1:GN , -c(1:2) ]
 wid2008 <- wid2008[ 1:GN , -c(1:2) ]
+
+# impute missing value
+output1995[1877] <- mltc23
+rm(mltc23)
 
 # save the workspace
 save.image(file = "data/TiVa.RData" )

### 2015-04-13 R/merge-wwz.R
merge from win and lin comps

@@ -17,8 +17,6 @@ w2008 <- cbind(year = 2008, w2008)
 # stack the data sets
 w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
 
-<<<<<<< HEAD
-
 # to lower case
 w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
 w1995_2008["Exporting_Indsutry"] <- tolower( w1995_2008["Exporting_Industry"] )
@@ -26,7 +24,3 @@ w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
 
 # save the data
 save(w1995_2008, file = "data/w1995_2008.RData")
-=======
-# save the data
-save(w1995_2008, file = "data/w1995_2008.RData" )
->>>>>>> 08d2e4f969dd117f45a58f120e274cad80214d03

### 2015-04-13 R/merge-wwz.R
update load data

@@ -3,10 +3,7 @@
 # bquast@gmail.com
 
 # load the data
-load("data/w1995.RData")
-load("data/w2000.RData")
-load("data/w2005.RData")
-load("data/w2008.RData")
+load("data/TiVa-WWZ.RData")
 
 # add year variable to each data set
 w1995 <- cbind(year = 1995, w1995)

### 2015-04-13 R/merge-wwz.R
fix typo

@@ -16,7 +16,7 @@ w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
 
 # to lower case
 w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
-w1995_2008["Exporting_Indsutry"] <- tolower( w1995_2008["Exporting_Industry"] )
+w1995_2008["Exporting_Industry"] <- tolower( w1995_2008["Exporting_Industry"] )
 w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
 
 # save the data

### 2015-04-14 R/merge-wwz.R
use right source files, remove saving since not working

@@ -3,7 +3,10 @@
 # bquast@gmail.com
 
 # load the data
-load("data/TiVa-WWZ.RData")
+load("data/w1995.RData")
+load("data/w2000.RData")
+load("data/w2005.RData")
+load("data/w2008.RData")
 
 # add year variable to each data set
 w1995 <- cbind(year = 1995, w1995)
@@ -18,6 +21,3 @@ w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
 w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
 w1995_2008["Exporting_Industry"] <- tolower( w1995_2008["Exporting_Industry"] )
 w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
-
-# save the data
-save(w1995_2008, file = "data/w1995_2008.RData")

### 2015-04-14 R/merge-wwz.R
remove tolower of factors, fix saving issue

@@ -17,7 +17,10 @@ w2008 <- cbind(year = 2008, w2008)
 # stack the data sets
 w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
 
-# to lower case
-w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
-w1995_2008["Exporting_Industry"] <- tolower( w1995_2008["Exporting_Industry"] )
-w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
+# # to lower case
+# w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
+# w1995_2008["Exporting_Industry"] <- tolower( w1995_2008["Exporting_Industry"] )
+# w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
+
+# save
+save(w1995_2008, file = "data/w1995_2008.RData")

### 2015-04-14 R/analyse_gvc_indicators.R
add plot with 2 lines

@@ -30,6 +30,14 @@ gvc_indicators %>%
   ggvis(~year, ~e2r ) %>%
   layer_lines()
 
+# both
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik), i2e = sum(fvax_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~e2r ) %>%
+  layer_lines(stroke="e2r") %>%
+  layer_lines(~year, ~i2e, stroke="i2e")
+
 # add PVC
 # add intermediate imports (?)
 

### 2015-04-14 R/merge-wwz.R
fix lower case

@@ -18,9 +18,12 @@ w2008 <- cbind(year = 2008, w2008)
 w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
 
 # # to lower case
-# w1995_2008["Exporting_Country"]  <- tolower( w1995_2008["Exporting_Country"] )
-# w1995_2008["Exporting_Industry"] <- tolower( w1995_2008["Exporting_Industry"] )
-# w1995_2008["Importing_Country"]  <- tolower( w1995_2008["Importing_Country"] )
+levels(w1995_2008$Exporting_Country) <- tolower(levels(w1995_2008$Exporting_Country))
+levels(w1995_2008$Exporting_Industry) <- tolower(levels(w1995_2008$Exporting_Industry))
+levels(w1995_2008$Importing_Country) <- tolower(levels(w1995_2008$Importing_Country))
+
+# merge
+coun_var <- read.csv("data/Country_variables.csv")
 
 # save
 save(w1995_2008, file = "data/w1995_2008.RData")

### 2015-04-14 R/analyse_gvc_indicators.R
double line plots

@@ -33,13 +33,11 @@ gvc_indicators %>%
 # both
 gvc_indicators %>%
   group_by(year) %>%
-  summarise(e2r = sum(dvar_ik) / sum(exports_ik), i2e = sum(fvax_ik) / sum(exports_ik) ) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik),
+            i2e = sum(fvax_ik) / sum(exports_ik) ) %>%
   ggvis(~year, ~e2r ) %>%
   layer_lines(stroke="e2r") %>%
   layer_lines(~year, ~i2e, stroke="i2e")
-
-# add PVC
-# add intermediate imports (?)
 
 # plot nrca
 nrca %>%
@@ -47,15 +45,29 @@ nrca %>%
   ggvis(~year, ~nrca) %>%
   layer_lines()
 
+# add PVC
+# add intermediate imports (?)
+
 # plot wwz
+# RDV
 w1995_2008 %>%
   group_by(year) %>%
   summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) %>%
   ggvis(~year, ~RDV) %>%
   layer_lines()
 
+# PDC
 w1995_2008 %>%
   group_by(year) %>%
   summarise( PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
   ggvis(~year, ~PDC) %>%
   layer_lines()
+
+# both
+w1995_2008 %>%
+  group_by(year) %>%
+  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
+             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
+  ggvis(~year, ~RDV, stroke="RDV") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~PDC, stroke="PDC")

### 2015-04-14 R/analyse_gvc_indicators.R
with rename

@@ -48,7 +48,7 @@ nrca %>%
 # add PVC
 # add intermediate imports (?)
 
-# plot wwz
+# plot gvc_length
 # RDV
 w1995_2008 %>%
   group_by(year) %>%
@@ -71,3 +71,5 @@ w1995_2008 %>%
   ggvis(~year, ~RDV, stroke="RDV") %>%
   layer_lines() %>%
   layer_lines(~year, ~PDC, stroke="PDC")
+
+#

### 2015-04-14 R/analyse_gvc_indicators.R
start of industry

@@ -73,3 +73,5 @@ w1995_2008 %>%
   layer_lines(~year, ~PDC, stroke="PDC")
 
 #
+w1995_2008 %>%
+  group_by(Exporting_Industry)

### 2015-04-15 R/merge-wwz.R
save as RDS because < 50mb

@@ -22,8 +22,11 @@ levels(w1995_2008$Exporting_Country) <- tolower(levels(w1995_2008$Exporting_Coun
 levels(w1995_2008$Exporting_Industry) <- tolower(levels(w1995_2008$Exporting_Industry))
 levels(w1995_2008$Importing_Country) <- tolower(levels(w1995_2008$Importing_Country))
 
-# merge
-coun_var <- read.csv("data/Country_variables.csv")
+# remove ROW (Rest of World)
+w1995_2008 <- subset(w1995_2008, Exporting_Country != "row")
 
 # save
-save(w1995_2008, file = "data/w1995_2008.RData")
+saveRDS(w1995_2008, file = "data/w1995_2008.rds", compress = "xz")
+
+# merge
+coun_var <- read.csv("data/Country_variables.csv")

### 2015-04-15 R/analyse_gvc_indicators.R
add plotting by industry

@@ -74,4 +74,6 @@ w1995_2008 %>%
 
 #
 w1995_2008 %>%
-  group_by(Exporting_Industry)
+  # group_by(year, Exporting_Industry) %>%
+  ggvis(~year, ~DVA_FIN) %>%
+  layer_lines()

### 2015-04-16 R/analyse_gvc_indicators.R
more updates

@@ -74,6 +74,7 @@ w1995_2008 %>%
 
 #
 w1995_2008 %>%
-  # group_by(year, Exporting_Industry) %>%
-  ggvis(~year, ~DVA_FIN) %>%
+  group_by(year, Exporting_Industry) %>%
+  summarise(DVA_FIN = sum(DVA_FIN)) %>%
+  ggvis(~year, ~DVA_FIN, stroke = ~Exporting_Industry) %>%
   layer_lines()

### 2015-04-16 R/analyse_gvc_indicators.R
load from RDS

@@ -5,7 +5,7 @@
 # load the data
 load("data/gvc_indicators.RData")
 load("data/nrca_df.RData")
-load("data/w1995_2008.RData")
+w1995_2008 <- readRDS("data/w1995_2008.rds")
 
 # load required libraries
 library(dplyr)

### 2015-04-16 R/import_country_vars.R
import country vars

@@ -1,0 +1,12 @@
+# import-country-vars.R
+# Bastiaan Quast
+# bquast@gmail.com
+
+# load library
+library(readr)
+
+# load the data
+country_vars <- read_csv(file = "data/Country_variables.csv")
+
+# save the data
+save(country_vars, file = "data/country_vars.RData")

### 2015-04-16 R/analyse_gvc_indicators.R
move to other gvc ind

@@ -39,6 +39,14 @@ gvc_indicators %>%
   layer_lines(stroke="e2r") %>%
   layer_lines(~year, ~i2e, stroke="i2e")
 
+# by industry
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10) %>%
+  ggvis(y = ~i2e) %>%
+  layer_bars()
+
 # plot nrca
 nrca %>%
   group_by() %>%
@@ -72,9 +80,4 @@ w1995_2008 %>%
   layer_lines() %>%
   layer_lines(~year, ~PDC, stroke="PDC")
 
-#
-w1995_2008 %>%
-  group_by(year, Exporting_Industry) %>%
-  summarise(DVA_FIN = sum(DVA_FIN)) %>%
-  ggvis(~year, ~DVA_FIN, stroke = ~Exporting_Industry) %>%
-  layer_lines()
+

### 2015-04-16 R/merge-wwz.R
move to seperate file

@@ -27,6 +27,3 @@ w1995_2008 <- subset(w1995_2008, Exporting_Country != "row")
 
 # save
 saveRDS(w1995_2008, file = "data/w1995_2008.rds", compress = "xz")
-
-# merge
-coun_var <- read.csv("data/Country_variables.csv")

### 2015-04-16 R/analyse_gvc_indicators.R
industry plot

@@ -43,9 +43,10 @@ gvc_indicators %>%
 gvc_indicators %>%
   group_by(ind_name) %>%
   summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
-  top_n(10) %>%
-  ggvis(y = ~i2e) %>%
-  layer_bars()
+  top_n(10 %>%
+  ggvis(~factor(ind_name), ~i2e) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
 
 # plot nrca
 nrca %>%

### 2015-04-16 R/analyse_gvc_indicators.R
add plots

@@ -5,11 +5,25 @@
 # load the data
 load("data/gvc_indicators.RData")
 load("data/nrca_df.RData")
+load("data/country_vars.RData")
 w1995_2008 <- readRDS("data/w1995_2008.rds")
 
 # load required libraries
 library(dplyr)
 library(ggvis)
+
+# merge wwz and country vars
+w1995_2008 <- merge(w1995_2008,
+                    country_vars,
+                    by.x = c("year", "Exporting_Country"),
+                    by.y = c("year", "ctry")              )
+
+# merge gvc indicators and country vars
+gvc_indicators <- merge(gvc_indicators, country_vars, by = c("ctry", "year"))
+
+# calculate gdp (total)
+gvc_indicators$gdp <- gvc_indicators$avg_gdppc * gvc_indicators$pop
+gvc_indicators$avg_gdp <- gvc_indicators$avg_gdppc * gvc_indicators$avg_pop
 
 # basic exploratory analysis
 summary(gvc_indicators)
@@ -39,14 +53,46 @@ gvc_indicators %>%
   layer_lines(stroke="e2r") %>%
   layer_lines(~year, ~i2e, stroke="i2e")
 
+# i2e against GDP
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), avg_gdp = mean(avg_gdp) ) %>%
+  ggvis(~avg_gdp, ~i2e ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+
+# i2e against natural resources
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
+  ggvis(~nat_res_exp_sha_t_3, ~i2e ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+
+# e2r against natural resources
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
+  ggvis(~nat_res_exp_sha_t_3, ~e2r ) %>%
+  layer_points() %>%
[…]

### 2015-04-16 R/analyse_gvc_indicators.R
fix with ams, add lmic

@@ -33,7 +33,7 @@ summary(gvc_indicators)
 # i2e
 gvc_indicators %>%
   group_by(year) %>%
-  summarise(i2e = sum(fvax_ik) / sum(exports_ik) ) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
   ggvis(~year, ~i2e ) %>%
   layer_lines()
 
@@ -44,11 +44,11 @@ gvc_indicators %>%
   ggvis(~year, ~e2r ) %>%
   layer_lines()
 
-# both
+# GVC growth
 gvc_indicators %>%
   group_by(year) %>%
   summarise(e2r = sum(dvar_ik) / sum(exports_ik),
-            i2e = sum(fvax_ik) / sum(exports_ik) ) %>%
+            i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
   ggvis(~year, ~e2r ) %>%
   layer_lines(stroke="e2r") %>%
   layer_lines(~year, ~i2e, stroke="i2e")
@@ -126,3 +126,14 @@ w1995_2008 %>%
   ggvis(~year, ~RDV, stroke="RDV") %>%
   layer_lines() %>%
   layer_lines(~year, ~PDC, stroke="PDC")
+
+### LMICs
+
+# GVC growth
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(e2r = sum(dvar_lomidinc_ik) / sum(exports_ik),
+            i2e = sum(fvax_lomidinc_ams_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~e2r ) %>%
+  layer_lines(stroke="e2r") %>%
+  layer_lines(~year, ~i2e, stroke="i2e")

### 2015-04-16 R/analyse_gvc_indicators.R
hi lo

@@ -132,8 +132,8 @@ w1995_2008 %>%
 # GVC growth
 gvc_indicators %>%
   group_by(year) %>%
-  summarise(e2r = sum(dvar_lomidinc_ik) / sum(exports_ik),
-            i2e = sum(fvax_lomidinc_ams_ik) / sum(exports_ik) ) %>%
-  ggvis(~year, ~e2r ) %>%
-  layer_lines(stroke="e2r") %>%
-  layer_lines(~year, ~i2e, stroke="i2e")
+  summarise(i2e_lomid = sum(fvax_lomidinc_ams_ik) / sum(exports_ik),
+            i2e_hi    = sum(fvax_hiinc_ams_ik)    / sum(exports_ik) ) %>%
+  ggvis(~year, ~i2e_lomid ) %>%
+  layer_lines(stroke="i2e_lomid") %>%
+  layer_lines(~year, ~i2e_hi, stroke="i2e_hi")

### 2015-04-19 R/analyse_gvc_indicators.R
add ic plots

@@ -184,19 +184,28 @@ gvc_indicators %>%
   layer_lines(~year, ~e2r_hi, stroke="e2r_hi")
 
 # LMIC defition III
-# i2e hi lo
+# i2e by year, ic
 gvc_indicators %>%
   group_by(year, ic) %>%
   summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
   ggvis(~year, ~i2e, stroke = ~ic ) %>%
   layer_lines()
 
-# e2r hi lo
+# i2e by country, ic
 gvc_indicators %>%
-  group_by(year, ic) %>%
-  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
-  ggvis(~year, ~e2r, stroke = ~ic ) %>%
-  layer_lines()
+  group_by(ctry, ic) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  ggvis(~factor(ctry), ~i2e, fill = ~ic ) %>%
+  layer_bars() %>%
+  add_axis("x", properties = axis_props(label=list(angle=45, align="left")) )
+
+# i2e by industry, ic
+gvc_indicators %>%
+  group_by(ind_name, ic) %>%
+  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
+  ggvis(~ind_name, ~i2e, fill = ~ic ) %>%
+  layer_bars() %>%
+  add_axis("x", properties = axis_props(label=list(angle=45, align="left")) )
 
 # i2e against GDP
 gvc_indicators %>%
@@ -210,5 +219,60 @@ gvc_indicators %>%
 gvc_indicators %>%
   group_by(ctry, ic) %>%
   summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
-  ggvis(~nat_res_exp_sha_t_3, ~i2e ) %>%
+  ggvis(~nat_res_exp_sha_t_3, ~i2e) %>%
   layer_points(fill = ~ic)
+
+# e2r by year, ic
+gvc_indicators %>%
+  group_by(year, ic) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~e2r, stroke = ~ic ) %>%
+  layer_lines()
+
[…]

### 2015-04-20 R/analyse_gvc_indicators.R
dva fin and dva int

@@ -12,6 +12,16 @@ w1995_2008 <- readRDS("data/w1995_2008.rds")
 library(dplyr)
 library(ggvis)
 
+# create factor gdp var
+country_vars$ic   <- ifelse(country_vars$avg_gdppc <= 6000, "lic", ifelse(country_vars$avg_gdppc > 12000, "hic", "mic")  )
+
+# redo in country vars
+# calculate gdp (total)
+gvc_indicators$gdp <- gvc_indicators$avg_gdppc * gvc_indicators$pop
+gvc_indicators$avg_gdp <- gvc_indicators$avg_gdppc * gvc_indicators$avg_pop
+# create mean of gdp across years
+# i.e. avg_pop*avg_gdppc
+
 # merge wwz and country vars
 w1995_2008 <- merge(w1995_2008,
                     country_vars,
@@ -21,26 +31,11 @@ w1995_2008 <- merge(w1995_2008,
 # merge gvc indicators and country vars
 gvc_indicators <- merge(gvc_indicators, country_vars, by = c("ctry", "year") )
 
-# merge gvc indicators and nrca
-gvc_indicators <- merge(gvc_indicators, nrca, by.x = c("ctry", "isic"), by.y = c("country", "industry") )
-
-# calculate gdp (total)
-gvc_indicators$gdp <- gvc_indicators$avg_gdppc * gvc_indicators$pop
-gvc_indicators$avg_gdp <- gvc_indicators$avg_gdppc * gvc_indicators$avg_pop
-
-# basic exploratory analysis
-summary(gvc_indicators)
-
-# create mean of gdp across years
-# i.e. avg_pop*avg_gdppc
-
 # create logical gdp var
 gvc_indicators$lic  <- gvc_indicators$avg_gdppc <=  6000
 gvc_indicators$lmic <- gvc_indicators$avg_gdppc <= 12000
 gvc_indicators$hic  <- gvc_indicators$avg_gdppc >= 12000
 
-# create factor gdp var
-gvc_indicators$ic   <- ifelse(gvc_indicators$avg_gdppc <= 6000, "lic", ifelse(gvc_indicators$avg_gdppc > 12000, "hic", "mic")  )
 
 ## create basic summaries
 
@@ -104,12 +99,18 @@ gvc_indicators %>%
[…]

### 2015-04-20 R/analyse_gvc_indicators.R
add avg_pop

@@ -12,6 +12,12 @@ w1995_2008 <- readRDS("data/w1995_2008.rds")
 library(dplyr)
 library(ggvis)
 
+# contruct avg_pop in country vars
+country_vars %<>%
+  group_by(ctry) %>%
+  summarise(avg_pop = mean(pop)) %>%
+  merge(country_vars, . , by = "ctry")
+
 # create factor gdp var
 country_vars$ic   <- ifelse(country_vars$avg_gdppc <= 6000, "lic", ifelse(country_vars$avg_gdppc > 12000, "hic", "mic")  )
 

### 2015-04-20 R/analyse_gvc_indicators.R
revamp avg_gdp & pop

@@ -12,21 +12,15 @@ w1995_2008 <- readRDS("data/w1995_2008.rds")
 library(dplyr)
 library(ggvis)
 
-# contruct avg_pop in country vars
+# contruct avg_pop and gdp
 country_vars %<>%
   group_by(ctry) %>%
   summarise(avg_pop = mean(pop)) %>%
   merge(country_vars, . , by = "ctry")
+country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
 
 # create factor gdp var
 country_vars$ic   <- ifelse(country_vars$avg_gdppc <= 6000, "lic", ifelse(country_vars$avg_gdppc > 12000, "hic", "mic")  )
-
-# redo in country vars
-# calculate gdp (total)
-gvc_indicators$gdp <- gvc_indicators$avg_gdppc * gvc_indicators$pop
-gvc_indicators$avg_gdp <- gvc_indicators$avg_gdppc * gvc_indicators$avg_pop
-# create mean of gdp across years
-# i.e. avg_pop*avg_gdppc
 
 # merge wwz and country vars
 w1995_2008 <- merge(w1995_2008,

### 2015-04-20 R/analyse_gvc_indicators.R
add FVA FIN and FVA INT

@@ -288,25 +288,95 @@ w1995_2008 %>%
   ggvis(~year, ~DVA_FIN, stroke = ~ic) %>%
   layer_lines()
 
-# DVA FIN for LIC
-w1995_2008 %>%
-  filter(ic == "lic") %>%
-  group_by(year, Exporting_Country) %>%
-  summarise( DVA_FIN = mean(DVA_FIN) ) %>%
-  ggvis(~year, ~DVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
-  layer_lines()
+# # DVA FIN for LIC
+# w1995_2008 %>%
+#   filter(ic == "lic") %>%
+#   group_by(year, Exporting_Country) %>%
+#   summarise( DVA_FIN = mean(DVA_FIN) ) %>%
+#   ggvis(~year, ~DVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
+#   layer_lines()
+#
+# # DVA INT by LIC
+# w1995_2008 %>%
+#   group_by(year, ic) %>%
+#   summarise( DVA_INT = mean(DVA_INT) ) %>%
+#   ggvis(~year, ~DVA_INT, stroke = ~ic) %>%
+#   layer_lines()
+#
+# # DVA INT for LIC
+# w1995_2008 %>%
+#   filter(ic == "lic") %>%
+#   group_by(year, Exporting_Country) %>%
+#   summarise( DVA_INT = mean(DVA_INT) ) %>%
+#   ggvis(~year, ~DVA_INT, stroke = ~factor(Exporting_Country) ) %>%
+#   layer_lines()
 
-# DVA INT by LIC
+# back to sums
+# FVA FIN by year, IC
 w1995_2008 %>%
   group_by(year, ic) %>%
-  summarise( DVA_INT = mean(DVA_INT) ) %>%
-  ggvis(~year, ~DVA_INT, stroke = ~ic) %>%
+  summarise( FVA_FIN = sum(OVA_FIN) + sum(MVA_FIN) ) %>%
+  ggvis(~year, ~FVA_FIN, stroke = ~ic) %>%
   layer_lines()
 
-# DVA INT for LIC
+# FVA INT by year, IC
+w1995_2008 %>%
+  group_by(year, ic) %>%
+  summarise( FVA_INT = sum(OVA_INT) + sum(MVA_INT) ) %>%
+  ggvis(~year, ~FVA_INT, stroke = ~ic) %>%
+  layer_lines()
+
[…]

### 2015-04-20 R/analyse_gvc_indicators.R
add DViX plots

@@ -374,9 +374,18 @@ w1995_2008 %>%
   ggvis(~year, ~FVA_INT, stroke = ~factor(Exporting_Country) ) %>%
   layer_lines()
 
-# FVA INT by year, IC
+# FVA FIN by year, Exporting Country, for LIC
 w1995_2008 %>%
-  group_by(year, ic) %>%
+  group_by(year, Exporting_Country) %>%
+  summarise( FVA_FIN = sum(OVA_FIN) + sum(MVA_FIN) ) %>%
+  ggvis(~year, ~FVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
+  layer_lines()
+
+# DYNAMIC!!!
+# FVA INT by year, Exporting Country
+w1995_2008 %>%
+  group_by(year, ic, Exporting_Country) %>%
   summarise( FVA_INT = sum(OVA_INT) + sum(MVA_INT) ) %>%
-  ggvis(~year, ~FVA_INT, stroke = ~ic) %>%
+  ggvis(~year, ~FVA_INT, stroke = ~Exporting_Country ) %>%
+  filter(ic == eval(input_radiobuttons(choices = c("lic", "mic", "hic"), selected = "lic") ) ) %>%
   layer_lines()

### 2015-04-21 R/analyse_gvc_indicators.R
add w1995-2008 since now < 50 mb

@@ -11,6 +11,11 @@ w1995_2008 <- readRDS("data/w1995_2008.rds")
 # load required libraries
 library(dplyr)
 library(ggvis)
+library(magrittr)
+
+# remove unneeded columns
+w1995_2008 %<>%
+  select(-(24:29))
 
 # contruct avg_pop and gdp
 country_vars %<>%

### 2015-04-21 R/analyse_gvc_indicators.R
load from new data file

@@ -6,7 +6,7 @@
 load("data/gvc_indicators.RData")
 load("data/nrca_df.RData")
 load("data/country_vars.RData")
-w1995_2008 <- readRDS("data/w1995_2008.rds")
+load("data/w1995_2008.RData")
 
 # load required libraries
 library(dplyr)

### 2015-04-21 R/analyse_gvc_indicators.R
move to merge file

@@ -13,10 +13,6 @@ library(dplyr)
 library(ggvis)
 library(magrittr)
 
-# remove unneeded columns
-w1995_2008 %<>%
-  select(-(24:29))
-
 # contruct avg_pop and gdp
 country_vars %<>%
   group_by(ctry) %>%

### 2015-04-21 R/merge-wwz.R

@@ -25,5 +25,9 @@ levels(w1995_2008$Importing_Country) <- tolower(levels(w1995_2008$Importing_Coun
 # remove ROW (Rest of World)
 w1995_2008 <- subset(w1995_2008, Exporting_Country != "row")
 
+# remove unneeded columns
+w1995_2008 %<>%
+  select(-(24:29))
+
 # save
-saveRDS(w1995_2008, file = "data/w1995_2008.rds", compress = "xz")
+save(w1995_2008, file = "data/w1995_2008.RData")

### 2015-04-21 R/analyse_gvc_indicators.R
also move

@@ -11,7 +11,6 @@ load("data/w1995_2008.RData")
 # load required libraries
 library(dplyr)
 library(ggvis)
-library(magrittr)
 
 # contruct avg_pop and gdp
 country_vars %<>%

### 2015-04-21 R/merge-wwz.R

@@ -2,6 +2,9 @@
 # Bastiaan Quast
 # bquast@gmail.com
 
+# load libraries
+library(magrittr)
+
 # load the data
 load("data/w1995.RData")
 load("data/w2000.RData")

### 2015-04-21 R/merge-wwz.R
use pipe

@@ -26,11 +26,10 @@ levels(w1995_2008$Exporting_Industry) <- tolower(levels(w1995_2008$Exporting_Ind
 levels(w1995_2008$Importing_Country) <- tolower(levels(w1995_2008$Importing_Country))
 
 # remove ROW (Rest of World)
-w1995_2008 <- subset(w1995_2008, Exporting_Country != "row")
+w1995_2008 %<>% subset(Exporting_Country != "row")
 
 # remove unneeded columns
-w1995_2008 %<>%
-  select(-(24:29))
+w1995_2008 %<>% select(-(24:29))
 
 # save
 save(w1995_2008, file = "data/w1995_2008.RData")

### 2015-04-21 R/analyse_gvc_indicators.R
merge using magrittr %<>%

@@ -11,25 +11,27 @@ load("data/w1995_2008.RData")
 # load required libraries
 library(dplyr)
 library(ggvis)
+library(magrittr)
 
 # contruct avg_pop and gdp
 country_vars %<>%
   group_by(ctry) %>%
   summarise(avg_pop = mean(pop)) %>%
   merge(country_vars, . , by = "ctry")
+
 country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
 
 # create factor gdp var
-country_vars$ic   <- ifelse(country_vars$avg_gdppc <= 6000, "lic", ifelse(country_vars$avg_gdppc > 12000, "hic", "mic")  )
+country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
+                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )
 
 # merge wwz and country vars
-w1995_2008 <- merge(w1995_2008,
-                    country_vars,
-                    by.x = c("year", "Exporting_Country"),
-                    by.y = c("year", "ctry")              )
+w1995_2008 %<>% merge(country_vars,
+                      by.x = c("year", "Exporting_Country"),
+                      by.y = c("year", "ctry")              )
 
 # merge gvc indicators and country vars
-gvc_indicators <- merge(gvc_indicators, country_vars, by = c("ctry", "year") )
+gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
 
 # create logical gdp var
 gvc_indicators$lic  <- gvc_indicators$avg_gdppc <=  6000

### 2015-04-21 R/analyse_gvc_indicators.R
normalise by exports

@@ -22,7 +22,7 @@ country_vars %<>%
 country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
 
 # create factor gdp var
-country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
+country_vars$ic      <- with( ifelse(avg_gdppc <= 6000, "lic",
                                                   ifelse(avg_gdppc > 12000, "hic", "mic") ) )
 
 # merge wwz and country vars
@@ -271,25 +271,25 @@ gvc_indicators %>%
 # RDV
 w1995_2008 %>%
   group_by(year, ic) %>%
-  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) %>%
+  summarise( RDV = ( sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) / sum(texp) ) %>%
   ggvis(~year, ~RDV, stroke = ~ic) %>%
   layer_lines()
 
 # PDC
 w1995_2008 %>%
   group_by(year, ic) %>%
-  summarise( PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
+  summarise( PDC = ( sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) / sum(texp) ) %>%
   ggvis(~year, ~PDC, stroke = ~ic) %>%
   layer_lines()
 
-# using mean here
-# DVA FIN by IC
-w1995_2008 %>%
-  group_by(year, ic) %>%
-  summarise( DVA_FIN = mean(DVA_FIN) ) %>%
-  ggvis(~year, ~DVA_FIN, stroke = ~ic) %>%
-  layer_lines()
-
+# # using mean here
+# # DVA FIN by IC
+# w1995_2008 %>%
+#   group_by(year, ic) %>%
+#   summarise( DVA_FIN = mean(DVA_FIN) ) %>%
+#   ggvis(~year, ~DVA_FIN, stroke = ~ic) %>%
+#   layer_lines()
+#
 # # DVA FIN for LIC
 # w1995_2008 %>%
 #   filter(ic == "lic") %>%

### 2015-04-21 R/analyse_gvc_indicators.R
normalise by exports

@@ -313,18 +313,17 @@ w1995_2008 %>%
 #   ggvis(~year, ~DVA_INT, stroke = ~factor(Exporting_Country) ) %>%
 #   layer_lines()
 
-# back to sums
 # FVA FIN by year, IC
 w1995_2008 %>%
   group_by(year, ic) %>%
-  summarise( FVA_FIN = sum(OVA_FIN) + sum(MVA_FIN) ) %>%
+  summarise( FVA_FIN = ( sum(OVA_FIN) + sum(MVA_FIN) ) / sum(texp) ) %>%
   ggvis(~year, ~FVA_FIN, stroke = ~ic) %>%
   layer_lines()
 
 # FVA INT by year, IC
 w1995_2008 %>%
   group_by(year, ic) %>%
-  summarise( FVA_INT = sum(OVA_INT) + sum(MVA_INT) ) %>%
+  summarise( FVA_INT = ( sum(OVA_INT) + sum(MVA_INT) ) / sum(texp) ) %>%
   ggvis(~year, ~FVA_INT, stroke = ~ic) %>%
   layer_lines()
 
@@ -332,7 +331,7 @@ w1995_2008 %>%
 w1995_2008 %>%
   filter(ic == "lic") %>%
   group_by(year, Exporting_Country) %>%
-  summarise( FVA_FIN = sum(OVA_FIN) + sum(MVA_FIN) ) %>%
+  summarise( FVA_FIN = ( sum(OVA_FIN) + sum(MVA_FIN) ) / sum(texp) ) %>%
   ggvis(~year, ~FVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
   layer_lines()
 
@@ -340,7 +339,7 @@ w1995_2008 %>%
 w1995_2008 %>%
   filter(ic == "lic") %>%
   group_by(year, Exporting_Country) %>%
-  summarise( FVA_INT = sum(OVA_INT) + sum(MVA_INT) ) %>%
+  summarise( FVA_INT = ( sum(OVA_INT) + sum(MVA_INT) ) / sum(texp) ) %>%
   ggvis(~year, ~FVA_INT, stroke = ~factor(Exporting_Country) ) %>%
   layer_lines()
 
@@ -348,7 +347,7 @@ w1995_2008 %>%
 w1995_2008 %>%
   filter(ic == "mic") %>%
   group_by(year, Exporting_Country) %>%
[…]

### 2015-04-21 R/analyse_gvc_indicators.R
remove double and normalise dynamic by exports

@@ -375,18 +375,11 @@ w1995_2008 %>%
   ggvis(~year, ~FVA_INT, stroke = ~factor(Exporting_Country) ) %>%
   layer_lines()
 
-# FVA FIN by year, Exporting Country, for LIC
-w1995_2008 %>%
-  group_by(year, Exporting_Country) %>%
-  summarise( FVA_FIN = sum(OVA_FIN) + sum(MVA_FIN) ) %>%
-  ggvis(~year, ~FVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
-  layer_lines()
-
 # DYNAMIC!!!
 # FVA INT by year, Exporting Country
 w1995_2008 %>%
   group_by(year, ic, Exporting_Country) %>%
-  summarise( FVA_INT = sum(OVA_INT) + sum(MVA_INT) ) %>%
+  summarise( FVA_INT = ( sum(OVA_INT) + sum(MVA_INT) ) / sum(texp) ) %>%
   ggvis(~year, ~FVA_INT, stroke = ~Exporting_Country ) %>%
   filter(ic == eval(input_radiobuttons(choices = c("lic", "mic", "hic"), selected = "lic") ) ) %>%
   layer_lines()

### 2015-04-21 R/analyse_gvc_indicators.R
add NRCA for LICs (aggregate) and china (industry)

@@ -33,10 +33,13 @@ w1995_2008 %<>% merge(country_vars,
 # merge gvc indicators and country vars
 gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
 
-# create logical gdp var
-gvc_indicators$lic  <- gvc_indicators$avg_gdppc <=  6000
-gvc_indicators$lmic <- gvc_indicators$avg_gdppc <= 12000
-gvc_indicators$hic  <- gvc_indicators$avg_gdppc >= 12000
+# merge gvc indicators and nrca
+gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
+
+# # create logical gdp var
+# gvc_indicators$lic  <- gvc_indicators$avg_gdppc <=  6000
+# gvc_indicators$lmic <- gvc_indicators$avg_gdppc <= 12000
+# gvc_indicators$hic  <- gvc_indicators$avg_gdppc >= 12000
 
 
 ## create basic summaries
@@ -105,17 +108,25 @@ gvc_indicators %>%
   layer_bars() %>%
   add_axis("x", title = "industry name")
 
-# plot nrca
-nrca %>%
-  group_by(country) %>%
-  ggvis(~year, ~nrca) %>%
-  layer_lines(stroke=~country)
+## plot nrca
 
-nrca %>%
-  filter(ic != "hic")
-  group_by(country) %>%
-  ggvis(~year, ~nrca) %>%
-  layer_lines(stroke=~country)
+# aggregate NRCA of LICs
+gvc_indicators %>%
+  filter(ic == "lic") %>%
+  group_by(ctry, year) %>%
+  summarise( nrca = sum(nrca) ) %>%
+  ggvis(~year, ~nrca, stroke=~factor(ctry) ) %>%
+  layer_lines()
+
+# NRCA china by industry
+gvc_indicators %>%
+  filter(ctry == "chn") %>%
+  group_by(year, ind_name) %>%
+  summarise( nrca = sum(nrca) ) %>%
+  ggvis(~year, ~nrca, stroke=~ind_name ) %>%
+  layer_lines()
+
+
 
 # add PVC
 # add intermediate imports (?)

### 2015-04-21 R/analyse_gvc_indicators.R
DViX for lic

@@ -91,7 +91,7 @@ gvc_indicators %>%
   layer_points() %>%
   layer_model_predictions(model = "lm")
 
-# by country
+# i2e by country
 gvc_indicators %>%
   group_by(ctry) %>%
   summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
@@ -99,7 +99,7 @@ gvc_indicators %>%
   ggvis(~factor(ctry), ~i2e ) %>%
   layer_bars()
 
-# by industry
+# i2e by industry
 gvc_indicators %>%
   group_by(ind_name) %>%
   summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
@@ -108,7 +108,8 @@ gvc_indicators %>%
   layer_bars() %>%
   add_axis("x", title = "industry name")
 
-## plot nrca
+
+## NRCA
 
 # aggregate NRCA of LICs
 gvc_indicators %>%
@@ -278,7 +279,6 @@ gvc_indicators %>%
 
 # WWZ by IC
 
-# use mean in stead of sum?
 # RDV
 w1995_2008 %>%
   group_by(year, ic) %>%
@@ -293,36 +293,8 @@ w1995_2008 %>%
   ggvis(~year, ~PDC, stroke = ~ic) %>%
   layer_lines()
 
-# # using mean here
-# # DVA FIN by IC
-# w1995_2008 %>%
-#   group_by(year, ic) %>%
-#   summarise( DVA_FIN = mean(DVA_FIN) ) %>%
-#   ggvis(~year, ~DVA_FIN, stroke = ~ic) %>%
-#   layer_lines()
-#
-# # DVA FIN for LIC
-# w1995_2008 %>%
-#   filter(ic == "lic") %>%
-#   group_by(year, Exporting_Country) %>%
-#   summarise( DVA_FIN = mean(DVA_FIN) ) %>%
-#   ggvis(~year, ~DVA_FIN, stroke = ~factor(Exporting_Country) ) %>%
-#   layer_lines()
-#
-# # DVA INT by LIC
-# w1995_2008 %>%
-#   group_by(year, ic) %>%
-#   summarise( DVA_INT = mean(DVA_INT) ) %>%
-#   ggvis(~year, ~DVA_INT, stroke = ~ic) %>%
-#   layer_lines()
-#
-# # DVA INT for LIC
-# w1995_2008 %>%
[…]

### 2015-04-21 OECD-TiVa-LMIC-GVCs.Rmd
add package description
adde xample graph

@@
 
+ ```{r}
+ library(decompr)
+ ls.str("package:decompr")
+ ```
+
+ ```{r}
+ library(gvc)
+ ls.str("package:gvc")
+ ```
+
+ ```{r, echo=FALSE}
+ # load the data
+load("data/gvc_indicators.RData")
+load("data/nrca_df.RData")
+load("data/country_vars.RData")
+load("data/w1995_2008.RData")
+
+ # load required libraries
+library(dplyr)
+library(ggvis)
+library(magrittr)
+
+# contruct avg_pop and gdp
+country_vars %<>%
+  group_by(ctry) %>%
+  summarise(avg_pop = mean(pop)) %>%
+  merge(country_vars, . , by = "ctry")
+
+country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
+
+# create factor gdp var
+country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
+                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )
+
+# merge wwz and country vars
+w1995_2008 %<>% merge(country_vars,
+                      by.x = c("year", "Exporting_Country"),
+                      by.y = c("year", "ctry")              )
+
+# merge gvc indicators and country vars
+gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
+
+# merge gvc indicators and nrca
+gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
+```
+
 ## General Stylised Facts
@@
  - trade balanses in VA terms differ substantially from those in gross terms
+
+ ```{r}
+ gvc_indicators %>%
+  group_by(year, ic) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~i2e, stroke = ~ic ) %>%
+  layer_lines()
+ ```
 

### 2015-04-21 R/analyse_gvc_indicators.R

@@ -22,7 +22,7 @@ country_vars %<>%
 country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
 
 # create factor gdp var
-country_vars$ic      <- with( ifelse(avg_gdppc <= 6000, "lic",
+country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
                                                   ifelse(avg_gdppc > 12000, "hic", "mic") ) )
 
 # merge wwz and country vars

### 2015-04-21 OECD-TiVa-LMIC-GVCs.Rmd
working with plots

@@
 
- ```{r}
+## I2E
+
+  ```{r, echo=FALSE}
  gvc_indicators %>%

### 2015-04-21 OECD-TiVa-LMIC-GVCs.Rmd
add trca nrca plot

@@
 load("data/w1995_2008.RData")
+load("data/trca.RData")
 

### 2015-04-21 R/analyse_gvc_indicators.R

@@ -7,6 +7,7 @@ load("data/gvc_indicators.RData")
 load("data/nrca_df.RData")
 load("data/country_vars.RData")
 load("data/w1995_2008.RData")
+load("data/trca.RData")
 
 # load required libraries
 library(dplyr)
@@ -35,6 +36,7 @@ gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
 
 # merge gvc indicators and nrca
 gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
+gvc_indicators %<>% merge(trca, by = c("ctry", "isic", "year") )
 
 # # create logical gdp var
 # gvc_indicators$lic  <- gvc_indicators$avg_gdppc <=  6000
@@ -109,7 +111,37 @@ gvc_indicators %>%
   add_axis("x", title = "industry name")
 
 
-## NRCA
+## TRCA NRCA
+
+# TRCA vs NRCA for China, ELQ
+gvc_indicators %>%
+  filter(ctry == "chn") %>%
+  filter(ind_name == "elq") %>%
+  group_by(year) %>%
+  summarise( trca = sum(rca.x), nrca = sum(nrca) ) %>%
+  ggvis(~year, ~trca, stroke = "trca") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~nrca, stroke="nrca")
+
+# TRCA vs NRCA for China: CEQ
+gvc_indicators %>%
+  filter(ctry == "chn") %>%
+  filter(ind_name == "ceq") %>%
+  group_by(year) %>%
+  summarise( trca = sum(rca.x), nrca = sum(nrca) ) %>%
+  ggvis(~year, ~trca, stroke = "trca") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~nrca, stroke="nrca")
+
+# TRCA vs NRCA for China: CEQ
+gvc_indicators %>%
+  filter(ctry == "chn") %>%
+  filter(ind_name == "ceq") %>%
+  group_by(year) %>%
+  summarise( trca = sum(rca.x), nrca = sum(nrca) ) %>%
+  ggvis(~year, ~trca, stroke = "trca") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~nrca, stroke="nrca")
 
 # aggregate NRCA of LICs
 gvc_indicators %>%

### 2015-04-21 OECD-TiVa-LMIC-GVCs.Rmd
add LaTeX
add CEQ for china plot to pres

@@
 ## This paper
+
+$$E_x$$
 
@@
 gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
+gvc_indicators %<>% merge(trca, by = c("ctry", "isic", "year") )
 ```
@@
 
+## CEQ of China
+
+```{r, echo=FALSE} 
+# TRCA vs NRCA for China, ELQ
+gvc_indicators %>%
+  filter(ctry == "chn") %>%
+  filter(ind_name == "elq") %>%
+  group_by(year) %>%
+  summarise( trca = sum(rca), nrca = sum(nrca) ) %>%
+  ggvis(~year, ~trca, stroke = "trca") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~nrca, stroke="nrca")
+```
+
 ## Conclusion

### 2015-04-21 R/analyse_gvc_indicators.R

@@ -118,7 +118,7 @@ gvc_indicators %>%
   filter(ctry == "chn") %>%
   filter(ind_name == "elq") %>%
   group_by(year) %>%
-  summarise( trca = sum(rca.x), nrca = sum(nrca) ) %>%
+  summarise( trca = sum(rca), nrca = sum(nrca) ) %>%
   ggvis(~year, ~trca, stroke = "trca") %>%
   layer_lines() %>%
   layer_lines(~year, ~nrca, stroke="nrca")
@@ -128,7 +128,7 @@ gvc_indicators %>%
   filter(ctry == "chn") %>%
   filter(ind_name == "ceq") %>%
   group_by(year) %>%
-  summarise( trca = sum(rca.x), nrca = sum(nrca) ) %>%
+  summarise( trca = sum(rca), nrca = sum(nrca) ) %>%
   ggvis(~year, ~trca, stroke = "trca") %>%
   layer_lines() %>%
   layer_lines(~year, ~nrca, stroke="nrca")

### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
Updated slieds

@@
 ---
-title: "OECD TiVa LMIC GVCs"
+title: "Global Value Chain Analysis of LMICs"
 author: "Victor Kummritz and Bastiaan Quast"
-date: "March 24, 2015"
+date: "April 22, 2015"
 output: ioslides_presentation
@@
 
-## Motivation
+## Motivation: The rise of GVCs
 
- - Global Value Chains relevance
- - GVCs in developement (capacity building)
- - General GVC picture overemphasises north-north because of higher monetary value
+* "[O]ne of the most important factors shaping contemporary international trade [is] the rise of global value chains."  
+*Pascal Lamy, September 2012*
+<br />  <br />
+* "We value [...] the relevance of regional and global value chains [...] in fostering economic growth, employment and development."  
+*G20 leaders, June 2012*
+<br /><br />
+* "Today, however, production processes increasingly involve global
+supply chains spanning multiple countries,[...]."  
+*Costinot, Vogel, and Wang, January 2013*
 
-## This paper
+## Motivation: GVCs and development
 
-$$E_x$$
+GVCs offer a new way to development: Countries can join value chains instead of building them:
 
-## Data
+  + World Economic Forum (2013): Reductions of GVC trade barriers
+could raise global GDP by 5\% and international trade by 15\%.
+  + UNCTAD (2013): Economies with the fastest growing GVC
+participation have GDP per capita growth rates around two percentage
+points above the average.
 
-ICIOs from TiVa (not WIOD)
+## Motivation: Current shortcomings in the GVCs analysis
 
-TiVa data
+ - Limited data availability reduces analysis to high- and middle-income countries, which might not be representative.
+
+ - Analysis so far mostly based on limited set of indicators.
+
+ - Technical complexity of standard GVC indicators represents a large entry barrier to GVC analysis.
+
+## This project
+
+- Develop tools that automate and simplify the derivation of GVC indicators.
+
+- Implement the calculation of novel GVC indicators.
+
+- Analyse these indicators using new data with extensive country coverage.
+
+## New Toolbox and Indicators: decompr and gvc
+Background: Leontief decomposition
+
+Applying the Leontief (1936) insight to Inter-Country Input-Output tables to derive value added multipliers ($VB$) following Hummels et al. (1998, 2001):
+$$
+VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...)
+$$
+$$
+= V (I-A)^{-1}.
+$$
+
+The multipliers can be post-multiplied with exports to estimate industry-level value added origins of each industry's exports.
+
+$V$ - Value added / Output, $A$ - Input-output coefficients.
+
+## New Toolbox and Indicators: decompr and gvc
+Exemplary indicators:
+
+- I2E (importing-to-export, based on Hummels et al. 1998): The foreign value added content of domestic exports.
+    - I2E_LMIC: The foreign value added content sourced in low- and middle income   countries
+    - etc.
+- E2R (exporting-to-reexport, based on Hummels et al. 1998): The domestic value added content of foreign exports.
+- NRCA (New revealed comparative advantage, based on Wang, Wei, and Zhu 2014): Revealed comparative advantage measure based on value added data.
+
+## An application to new data: ICIOS covering LMICs
+
+OECD ICIOs:
+
 - 57 countries
+
 - 34 sectors
+
+- 1995, 2000, 2005, 2008, 2009
+
 - LMICs
 
-## New Toolbox and Indicators
-
- - source / using countries specific indicators
- - decompor and gvc functions
+## An application to new data: Confirming stylized facts
 
@@
 
+## An application to new data: Confirming stylized facts
+The growth of GVC participation.
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik),
+            i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~e2r ) %>%
+  layer_lines(stroke="e2r") %>%
+  layer_lines(~year, ~i2e, stroke="i2e")
+```
+
+## An application to new data: Confirming stylized facts
+The expansion of GVCs.
+```{r, echo=FALSE}
+w1995_2008 %>%
+  group_by(year) %>%
+  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
+             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
+  ggvis(~year, ~RDV, stroke="RDV") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~PDC, stroke="PDC")
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10) %>%
+  ggvis(~factor(ind_name), ~i2e) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  summarise(e2r = sum(dvar_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10, e2r) %>%
+  ggvis(~factor(ind_name), ~e2r) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  top_n(10, i2e) %>%
+  ggvis(~factor(ctry), ~i2e ) %>%
+  layer_bars()
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
+  top_n(10, e2r) %>%
+  ggvis(~factor(ctry), ~e2r ) %>%
+  layer_bars()
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), avg_gdp = mean(avg_gdp) ) %>%
+  ggvis(~avg_gdp, ~i2e ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
+  ggvis(~nat_res_exp_sha_t_3, ~i2e ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+```
+
+## An application to new data: Confirming stylized facts
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
+  ggvis(~nat_res_exp_sha_t_3, ~e2r ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+```
+
 ## General Stylised Facts

### 2015-04-22 R/analyse_gvc_indicators.R
remove old definitions

@@ -189,49 +189,8 @@ w1995_2008 %>%
   layer_lines(~year, ~PDC, stroke="PDC")
 
 
-### LMICs
-# LMIC definition I
+## by *IC (LIC/MIC/HIC)
 
-# i2e hi lo
-gvc_indicators %>%
-  group_by(year) %>%
-  summarise(i2e_lomid = sum(fvax_lomidinc_ams_ik) / sum(exports_ik),
-            i2e_hi    = sum(fvax_hiinc_ams_ik)    / sum(exports_ik) ) %>%
-  ggvis(~year, ~i2e_lomid ) %>%
-  layer_lines(stroke="i2e_lomid") %>%
-  layer_lines(~year, ~i2e_hi, stroke="i2e_hi")
-
-# e2r hi lo
-gvc_indicators %>%
-  group_by(year) %>%
-  summarise(e2r_lomid = sum(dvar_lomidinc_ik) / sum(exports_ik),
-            e2r_hi    = sum(dvar_hiinc_ik)    / sum(exports_ik) ) %>%
-  ggvis(~year, ~e2r_lomid ) %>%
-  layer_lines(stroke="e2r_lomid") %>%
-  layer_lines(~year, ~e2r_hi, stroke="e2r_hi")
-
-
-# LMIC definition II
-
-# i2e hi lo
-gvc_indicators %>%
-  group_by(year) %>%
-  summarise(i2e_lomid = sum(fvax_lomidinc1_ams_ik) / sum(exports_ik),
-            i2e_hi    = sum(fvax_hiinc1_ams_ik)    / sum(exports_ik) ) %>%
-  ggvis(~year, ~i2e_lomid ) %>%
-  layer_lines(stroke="i2e_lomid") %>%
-  layer_lines(~year, ~i2e_hi, stroke="i2e_hi")
-
-# e2r hi lo
-gvc_indicators %>%
-  group_by(year) %>%
-  summarise(e2r_lomid = sum(dvar_lomidinc1_ik) / sum(exports_ik),
-            e2r_hi    = sum(dvar_hiinc1_ik)    / sum(exports_ik) ) %>%
-  ggvis(~year, ~e2r_lomid ) %>%
-  layer_lines(stroke="e2r_lomid") %>%
-  layer_lines(~year, ~e2r_hi, stroke="e2r_hi")
-
-# LMIC defition III
 # i2e by year, ic
 gvc_indicators %>%
[…]

### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
reorder slides

@@
 date: "April 22, 2015"
-output: ioslides_presentation
+output:
+  ioslides_presentation:
+    fig_height: 3
+    fig_retina: null
+    fig_width: 5
 ---
@@
 ## New Toolbox and Indicators: decompr and gvc
-Exemplary indicators:
+
+Indicators:
 
 - I2E (importing-to-export, based on Hummels et al. 1998): The foreign value added content of domestic exports.
-    - I2E_LMIC: The foreign value added content sourced in low- and middle income   countries
-    - etc.
 - E2R (exporting-to-reexport, based on Hummels et al. 1998): The domestic value added content of foreign exports.
@@
 
-## An application to new data: ICIOS covering LMICs
+## An application to new data: ICIOs covering LMICs
 
@@
 
-- LMICs
+- Low- and Middle-Income Countries (LMICs)
 
-## An application to new data: Confirming stylized facts
+## package: decompr
 
@@
  ```
+
+## package: gvc
 
@@
 
-## An application to new data: Confirming stylized facts
-The growth of GVC participation.
+## The expansion of GVCs
+
+```{r, echo=FALSE}
+w1995_2008 %>%
+  group_by(year) %>%
+  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
+             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
+  ggvis(~year, ~RDV, stroke="RDV") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~PDC, stroke="PDC")
+```
+
+## The growth of GVC participation.
+
 ```{r, echo=FALSE}
@@
 
-## An application to new data: Confirming stylized facts
-The expansion of GVCs.
+## GVC participation by *IC: I2E
+Importing to Export (I2E)
+
 ```{r, echo=FALSE}
-w1995_2008 %>%
-  group_by(year) %>%
-  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
-             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
-  ggvis(~year, ~RDV, stroke="RDV") %>%
-  layer_lines() %>%
-  layer_lines(~year, ~PDC, stroke="PDC")
+# i2e by year, ic
+gvc_indicators %>%
+  group_by(year, ic) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~i2e, stroke = ~ic ) %>%
+  layer_lines()
+```
+
+## GVC participation by *IC: E2R
+Exporting to Re-export (E2R)
+
+```{r, echo=FALSE}
+# e2r by year, ic
+gvc_indicators %>%
+  group_by(year, ic) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
+  ggvis(~year, ~e2r, stroke = ~ic ) %>%
+  layer_lines()
 ```
@@
 
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ind_name) %>%
-  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
-  top_n(10) %>%
-  ggvis(~factor(ind_name), ~i2e) %>%
-  layer_bars() %>%
-  add_axis("x", title = "industry name")
-```
-
-## An application to new data: Confirming stylized facts
-
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ind_name) %>%
-  summarise(e2r = sum(dvar_ik) * 100 / sum(exports_ik) ) %>%
-  top_n(10, e2r) %>%
-  ggvis(~factor(ind_name), ~e2r) %>%
-  layer_bars() %>%
-  add_axis("x", title = "industry name")
-```
-
-## An application to new data: Confirming stylized facts
-
-```{r, echo=FALSE}
+```{r, echo=FALSE, }
 gvc_indicators %>%
@@
 
-## General Stylised Facts
-
- - section 3.1 of wwz
- - growth of GVC trade
- - length of GVC
- - determinants of an industry's participation
- - determinants of a country's participation
- - trade balanses in VA terms differ substantially from those in gross terms
-
-## I2E
-
-  ```{r, echo=FALSE}
- gvc_indicators %>%
-  group_by(year, ic) %>%
-  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
-  ggvis(~year, ~i2e, stroke = ~ic ) %>%
-  layer_lines()
- ```
-
-## GVCs of LMICs
-
- - stylised facts for LMICs
- - source / using countries specific indicators for LMICs
- - nrca and trca of LMIC
- - fvafin vs. fvaint for LMICs (WWZ 3.3)
- - dvix decomposition (WWZ 3.4)
-
 ## CEQ of China
 
-```{r, echo=FALSE} 
+```{r, echo=FALSE, } 
 # TRCA vs NRCA for China, ELQ
@@
   filter(ctry == "chn") %>%
-  filter(ind_name == "elq") %>%
+  filter(ind_name == "ceq") %>%
   group_by(year) %>%

### 2015-04-22 R/analyse_gvc_indicators.R
sourcing

@@ -53,6 +53,17 @@ gvc_indicators %>%
   ggvis(~year, ~i2e ) %>%
   layer_lines()
 
+# i2e lo mi sourcing
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(i2e_lic = sum(fvax_loinc_ams_ik) / sum(exports_ik),
+            i2e_mic = sum(fvax_midinc_ams_ik) / sum(exports_ik),
+            i2e_hic = sum(fvax_hiinc_ams_ik) / sum(exports_ik)) %>%
+  ggvis(~year, ~i2e_lic, stroke = "i2e_lic" ) %>%
+  layer_lines() %>%
+  layer_lines(~year, ~i2e_mic, stroke="i2e_mic") %>%
+  layer_lines(~year, ~i2e_hic, stroke="i2e_hic")
+
 # e2r
 gvc_indicators %>%
   group_by(year) %>%
@@ -110,6 +121,15 @@ gvc_indicators %>%
   layer_bars() %>%
   add_axis("x", title = "industry name")
 
+# e2r by industry
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  filter(ind_name != "pvh" & ind_name != "edu" & ind_name != "egw" & ind_name != "con" & ind_name != "rea" & ind_name != "gov" & ind_name != "hth" & ind_name != "ptl") %>%
+  summarise(e2r = sum(dvar_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10) %>%
+  ggvis(~factor(ind_name), ~e2r) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
 
 ## TRCA NRCA
 
@@ -168,22 +188,22 @@ gvc_indicators %>%
 # RDV
 w1995_2008 %>%
   group_by(year) %>%
-  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) %>%
+  summarise( RDV = () sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) / sum(texp) ) %>%
   ggvis(~year, ~RDV) %>%
   layer_lines()
 
 # PDC
 w1995_2008 %>%
   group_by(year) %>%
-  summarise( PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
[…]

### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
finish presentation

@@
 ---
+
+ ```{r, echo=FALSE, warnings=FALSE}
+ # load the data
+load("data/gvc_indicators.RData")
+load("data/nrca_df.RData")
+load("data/country_vars.RData")
+load("data/w1995_2008.RData")
+load("data/trca.RData")
+
+ # load required libraries
+library(dplyr)
+library(ggvis)
+library(magrittr)
+
+# contruct avg_pop and gdp
+country_vars %<>%
+  group_by(ctry) %>%
+  summarise(avg_pop = mean(pop)) %>%
+  merge(country_vars, . , by = "ctry")
+
+country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
+
+# create factor gdp var
+country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
+                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )
+
+# merge wwz and country vars
+w1995_2008 %<>% merge(country_vars,
+                      by.x = c("year", "Exporting_Country"),
+                      by.y = c("year", "ctry")              )
+
+# merge gvc indicators and country vars
+gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
+
+# merge gvc indicators and nrca
+gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
+gvc_indicators %<>% merge(trca, by = c("ctry", "isic", "year") )
+```
 
@@
 
- ```{r, echo=FALSE}
- # load the data
-load("data/gvc_indicators.RData")
-load("data/nrca_df.RData")
-load("data/country_vars.RData")
-load("data/w1995_2008.RData")
-load("data/trca.RData")
+## The growth of GVC participation.
 
- # load required libraries
-library(dplyr)
-library(ggvis)
-library(magrittr)
-
-# contruct avg_pop and gdp
-country_vars %<>%
-  group_by(ctry) %>%
-  summarise(avg_pop = mean(pop)) %>%
-  merge(country_vars, . , by = "ctry")
-
-country_vars$avg_gdp <- with(country_vars, avg_pop * avg_gdppc)
-
-# create factor gdp var
-country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
-                                                  ifelse(avg_gdppc > 12000, "hic", "mic") ) )
-
-# merge wwz and country vars
-w1995_2008 %<>% merge(country_vars,
-                      by.x = c("year", "Exporting_Country"),
-                      by.y = c("year", "ctry")              )
-
-# merge gvc indicators and country vars
-gvc_indicators %<>% merge(country_vars, by = c("ctry", "year") )
-
-# merge gvc indicators and nrca
-gvc_indicators %<>% merge(nrca, by.x = c("ctry", "isic", "year"), by.y = c("country", "industry", "year") )
-gvc_indicators %<>% merge(trca, by = c("ctry", "isic", "year") )
-```
-
-## The expansion of GVCs
-
-```{r, echo=FALSE}
-w1995_2008 %>%
-  group_by(year) %>%
-  summarise( RDV = sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2),
-             PDC = sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) %>%
-  ggvis(~year, ~RDV, stroke="RDV") %>%
-  layer_lines() %>%
-  layer_lines(~year, ~PDC, stroke="PDC")
-```
-
-## The growth of GVC participation.
+Rising value added in 3rd country exports.
 
@@
 
+## The expansion of GVCs
+
+Increasing amounts of double counting indicate elongation of GVCs.
+
+```{r, echo=FALSE}
+w1995_2008 %>%
+  group_by(year) %>%
+  summarise( PDC = ( sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) / sum(texp) ) %>%
+  ggvis(~year, ~PDC) %>%
+  layer_lines()
+```
+
+## I2E by industry
+
+Distance to final demand predicts an industry's GVC participation.
+
+```{r, echo=FALSE}
+# i2e by industry
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  summarise(i2e = sum(fvax_ams_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10) %>%
+  ggvis(~factor(ind_name), ~i2e) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
+```
+
+## E2R by industry
+
+```{r, echo=FALSE}
+# e2r by industry
+gvc_indicators %>%
+  group_by(ind_name) %>%
+  filter(ind_name != "pvh" & ind_name != "edu" & ind_name != "egw" & ind_name != "con" & ind_name != "rea" & ind_name != "gov" & ind_name != "hth" & ind_name != "ptl") %>%
+  summarise(e2r = sum(dvar_ik) * 100 / sum(exports_ik) ) %>%
+  top_n(10) %>%
+  ggvis(~factor(ind_name), ~e2r) %>%
+  layer_bars() %>%
+  add_axis("x", title = "industry name")
+```
+
+## I2E by country
+
+country size and export composition determine GVC participation.
+
+```{r, echo=FALSE, }
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
+  top_n(10, i2e) %>%
+  ggvis(~factor(ctry), ~i2e ) %>%
+  layer_bars()
+```
+
+## E2R by country
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
+  top_n(10, e2r) %>%
+  ggvis(~factor(ctry), ~e2r ) %>%
+  layer_bars()
+```
+
+## Natural Resources and E2R
+
+```{r, echo=FALSE}
+gvc_indicators %>%
+  group_by(ctry) %>%
+  summarise(e2r = sum(dvar_ik) / sum(exports_ik), nat_res_exports = mean(nat_res_exp_sha_t_3) ) %>%
+  ggvis(~nat_res_exports, ~e2r ) %>%
+  layer_points() %>%
+  layer_model_predictions(model = "lm")
+```
+
+# GVC indicators for LMICs
+
 ## GVC participation by *IC: I2E
-Importing to Export (I2E)
+
+LMICs drive recent growth in GVC participation.
 
@@
 
+## Sourcing I2E
+
+```{r, echo=FALSE}
+# i2e lo mi sourcing
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(i2e_lic = sum(fvax_loinc_ams_ik) / sum(exports_ik),
+            i2e_mic = sum(fvax_midinc_ams_ik) / sum(exports_ik),
+            i2e_hic = sum(fvax_hiinc_ams_ik) / sum(exports_ik)) %>%
+  ggvis(~year, ~i2e_lic, stroke = "i2e_lic" ) %>%
+  layer_lines() %>%
+  layer_lines(~year, ~i2e_mic, stroke="i2e_mic") %>%
+  layer_lines(~year, ~i2e_hic, stroke="i2e_hic")
+```
+
 ## GVC participation by *IC: E2R
-Exporting to Re-export (E2R)
 
@@
 
-## An application to new data: Confirming stylized facts
+## Sourcing E2R
 
-```{r, echo=FALSE, }
+```{r, echo=FALSE}
+# e2r lo mi sourcing
 gvc_indicators %>%
-  group_by(ctry) %>%
-  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik) ) %>%
-  top_n(10, i2e) %>%
-  ggvis(~factor(ctry), ~i2e ) %>%
-  layer_bars()
+  group_by(year) %>%
+  summarise(e2r_lic = sum(dvar_loinc_ik) / sum(exports_ik),
+            e2r_mic = sum(dvar_midinc_ik) / sum(exports_ik),
+            e2r_hic = sum(dvar_hiinc_ik) / sum(exports_ik)) %>%
+  ggvis(~year, ~e2r_lic, stroke = "e2r_lic" ) %>%
+  layer_lines() %>%
+  layer_lines(~year, ~e2r_mic, stroke="e2r_mic") %>%
+  layer_lines(~year, ~e2r_hic, stroke="e2r_hic")
 ```
 
-## An application to new data: Confirming stylized facts
+# New Indicators for GVC Analysis
 
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ctry) %>%
-  summarise(e2r = sum(dvar_ik) / sum(exports_ik) ) %>%
-  top_n(10, e2r) %>%
-  ggvis(~factor(ctry), ~e2r ) %>%
-  layer_bars()
-```
+## NRCA
 
-## An application to new data: Confirming stylized facts
-
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ctry) %>%
-  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), avg_gdp = mean(avg_gdp) ) %>%
-  ggvis(~avg_gdp, ~i2e ) %>%
-  layer_points() %>%
-  layer_model_predictions(model = "lm")
-```
-
-## An application to new data: Confirming stylized facts
-
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ctry) %>%
-  summarise(i2e = sum(fvax_ams_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
-  ggvis(~nat_res_exp_sha_t_3, ~i2e ) %>%
-  layer_points() %>%
-  layer_model_predictions(model = "lm")
-```
-
-## An application to new data: Confirming stylized facts
-
-```{r, echo=FALSE}
-gvc_indicators %>%
-  group_by(ctry) %>%
-  summarise(e2r = sum(dvar_ik) / sum(exports_ik), nat_res_exp_sha_t_3 = mean(nat_res_exp_sha_t_3) ) %>%
-  ggvis(~nat_res_exp_sha_t_3, ~e2r ) %>%
-  layer_points() %>%
-  layer_model_predictions(model = "lm")
-```
-
-## CEQ of China
+Computer Equipment in China: Traditional overestimation
 
@@
 
+## NRCA II
+
+Motor and Transport Equipment: Traditional underestimation
+
+```{r, echo=FALSE, } 
+# TRCA vs NRCA for China, ELQ
+gvc_indicators %>%
+  filter(ctry == "tha") %>%
+  filter(ind_name == "mtr") %>%
+  group_by(year) %>%
+  summarise( trca = sum(rca), nrca = sum(nrca) ) %>%
+  ggvis(~year, ~trca, stroke = "trca") %>%
+  layer_lines() %>%
+  layer_lines(~year, ~nrca, stroke="nrca")
+```
+
+## I2E: Final Goods
+
+Final-good assembly is relatively low-skilled.
+
+```{r, echo=FALSE, warning=FALSE}
+# FVA FIN china thailand india indonesia
+w1995_2008 %>%
+  filter(Exporting_Country == "chn" | Exporting_Country == "tha" | Exporting_Country == "ind" | Exporting_Country == "idn") %>%
+  group_by(year, Exporting_Country) %>%
+  summarise( I2E_FIN = ( sum(OVA_FIN) + sum(MVA_FIN) ) / sum(texp) ) %>%
+  ggvis(~year, ~I2E_FIN, stroke = ~factor(Exporting_Country) ) %>%
+  layer_lines()
+```
+
+## I2E: Intermediate Goods
+
+Rising intermediate good exports indicate moving up the value chain.
+
+```{r, echo=FALSE, warning=FALSE}
+# FVA INT china thailand india indonesia
+w1995_2008 %>%
+  filter(Exporting_Country == "chn" | Exporting_Country == "tha" | Exporting_Country == "ind" | Exporting_Country == "idn") %>%
+  group_by(year, Exporting_Country) %>%
+  summarise( I2E_INT = ( sum(OVA_INT) + sum(MVA_INT) ) / sum(texp) ) %>%
+  ggvis(~year, ~I2E_INT, stroke = ~factor(Exporting_Country) ) %>%
+  layer_lines()
+```
+
 ## Conclusion
+
+  - new data
+  - focus on LMICs
+
+  - new indicators suited for non-traditional trade patterns
+  - non-trivial linear algebra involved
+  - simplified application of mew algorithms
+  - many more indicators and tools available in the packages
+    - decompr: decomposition
+    - gvc: analysis

### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
add website and thank you
remove notices

@@
 
- ```{r}
+ ```{r, tidy=FALSE}
  library(decompr)
@@
 
- ```{r}
+ ```{r, tidy=FALSE}
  library(gvc)
@@
 
-```{r, echo=FALSE}
+```{r echo=FALSE}
 w1995_2008 %>%
@@
 
-```{r, echo=FALSE}
+```{r echo=FALSE, warning=FALSE, message=FALSE}
 # i2e by industry
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 # e2r by industry
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 gvc_indicators %>%
@@
     - gvc: analysis
+
+  - http://qua.st/decompr
+
+# Thank You!

### 2015-04-22 OECD-TiVa-LMIC-GVCs.Rmd
final version

@@
 
- ```{r, echo=FALSE, warnings=FALSE}
+ ```{r, echo=FALSE, warnings=FALSE, message=FALSE}
  # load the data
@@
 
-## New Toolbox and Indicators: decompr and gvc
+## package: decompr
+
+Decompositions
+
+ - Leontief (post: exports / output)
+ - new Wang-Wei-Zhu algorithm
+
+ - citation info
+
+ ```{r, tidy=FALSE}
+ library(decompr)
+ ```
+
+## package: gvc
 
@@
  - NRCA (New revealed comparative advantage, based on Wang, Wei, and Zhu 2014): Revealed comparative advantage measure based on value added data.
+
+ - no citation info (yet!)
 
@@
 
-## package: decompr
-
- ```{r, tidy=FALSE}
- library(decompr)
- ls.str("package:decompr")
- ```
-
-## package: gvc
-
- ```{r, tidy=FALSE}
- library(gvc)
- ls.str("package:gvc")
- ```
+# Trends in GVCs
 
@@
 
-Motor and Transport Equipment: Traditional underestimation
+Motor and Transport Equipment in Thailand: Traditional underestimation
 
@@
   - non-trivial linear algebra involved
-  - simplified application of mew algorithms
+  - simplified application of new algorithms
   - many more indicators and tools available in the packages

### 2015-05-22 Intro.tex
Draft intro
First go at intro for data paper

\documentclass[11pt,a4paper]{article}
\usepackage{natbib}

\linespread{1.5}

\begin{document}

The emergence of Global Value Chains (GVCs) offers a new path to industrialisation for developing countries. As \cite{riba12} phrases it, internationally fragmented production allows developing countries to join existing supply chains instead of building them. This brings about many potential advantages for these countries. Connecting with firms from advanced nations allows developing nations, for instance, to benefit from their sophisticated technologies and know-how. In addition, relying on an existing production network frees them from constraints imposed by economies of scale and the increased specialisation, that GVCs imply, limits the negative impact of unproductive parts of the domestic supply chain. After all, when competition moves from goods to tasks, comparative advantage becomes much finer and does not require a broad range of productive stages domestically. Conditional evidence for such a positive impact of GVC participation in low- and middle-income countries is presented in \citet{viku15} and \citet{unct13}.

The quick expansion of GVCs has been documented in recent work. For instance, \citet{dahuetal98, dahuetal01} show in two early seminal contributions that GVCs are responsible for a major share of the total growth in world trade from 1970 to 1990. \citet{rojoguno12} and \citet{ribajalo13} find that this growth in GVC trade has even accelerated in the recent two decades. Furthermore, this work has not only revealed a rapid rise in production fragmentaion across borders but it has also re-evaluated important indicators of trade, such as bilateral trade imbalances and revealed comparative advantage\footnote{See \citeauthor{joamsoca15} (forthcoming) for an extensive review of the literature on GVCs and outsourcing.} showing that calculating value added trade indicators is central to a better understand of countries' trade patters and competitiveness.

However, these contributions typically have two shortcomings. Firstly, most evidence is based on data from high-income countries. The reason is that reliable time-series of both national and international input-output tables have only been available for these countries. In addition, the evidence is regularly based on a small sample of GVC indicators that hide valuable information stemming from more decomposed and disaggregated indicators.

In this paper we address these issues by applying a novel and more detailed gross export decomposition developed by \citet{zhwaetal13} to a new set of Inter-Country Input-Output tables (ICIOs) with extensive country coverage provided by the OECD. The new ICIOs allow us to get a better understanding of the GVC activities of low- and middle-income countries while the new decomposition allows us to zoom in more closely at these activities revealing information not available from standard GVC indicators.

\end{document}

### 2016-01-27 R/import.R
new instructions

@@ -1,6 +1,79 @@
 # import.R
+# ---------------------------
+# import data from csv files
+#
 # Bastiaan Quast
 # bquast@gmail.com
+
+
+# unzip
+unzip(zipfile = 'data/tiva.zip', exdir = 'data')
+
+
+# load data
+countries  <- read.csv(file = 'data/countries1995.csv',  header = FALSE, sep = '', stringsAsFactors = FALSE)
+industries <- read.csv(file = 'data/industries1995.csv', header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter1995  <- read.csv(file = 'data/inter1995.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2000  <- read.csv(file = 'data/inter2000.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2005  <- read.csv(file = 'data/inter2005.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2008  <- read.csv(file = 'data/inter2008.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2009  <- read.csv(file = 'data/inter2009.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2010  <- read.csv(file = 'data/inter2010.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+inter2011  <- read.csv(file = 'data/inter2011.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final1995  <- read.csv(file = 'data/final1995.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2000  <- read.csv(file = 'data/final2000.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2005  <- read.csv(file = 'data/final2005.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2008  <- read.csv(file = 'data/final2008.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2009  <- read.csv(file = 'data/final2009.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2010  <- read.csv(file = 'data/final2010.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
+final2011  <- read.csv(file = 'data/final2011.csv',      header = FALSE, sep = '', stringsAsFactors = FALSE)
[…]

### 2016-01-27 R/decompose.R
use new names for data and add new years

@@ -6,29 +6,47 @@
 library(decompr)
 
 # run the Leontief decomposition
-l1995 <- decomp(wid1995,
-                wfd1995,
+l1995 <- decomp(inter1995,
+                final1995,
                 countries,
-                sectors,
+                industries,
                 output1995,
                 method = "leontief")
-l2000 <- decomp(wid2000,
-                wfd2000,
+l2000 <- decomp(inter2000,
+                final2000,
                 countries,
-                sectors,
+                industries,
                 output2000,
                 method = "leontief")
-l2005 <- decomp(wid2005,
-                wfd2005,
+l2005 <- decomp(inter2005,
+                final2005,
                 countries,
-                sectors,
+                industries,
                 output2005,
                 method = "leontief")
-l2008 <- decomp(wid2008,
-                wfd2008,
+l2008 <- decomp(inter2008,
+                final2008,
                 countries,
-                sectors,
+                industries,
                 output2008,
+                method = "leontief")
+l2009 <- decomp(inter2009,
+                final2009,
+                countries,
+                industries,
+                output2009,
+                method = "leontief")
+l2010 <- decomp(inter2010,
+                final2010,
+                countries,
+                industries,
+                output2010,
+                method = "leontief")
+l2011 <- decomp(inter2011,
+                final2011,
+                countries,
+                industries,
+                output2011,
                 method = "leontief")
 
 # save Leontief the data
@@ -36,29 +54,29 @@ save(l1995, l2000, l2005, l2008, file = "data/TiVa-Leontief.RData")
 
 
 
-# run the Wang-Wei-Zhu decomposition 
-w1995 <- decomp(wid1995,
-                wfd1995,
+# run the Wang-Wei-Zhu decomposition
+w1995 <- decomp(inter1995,
+                final1995,
                 countries,
-                sectors,
+                industries,
                 output1995,
                 method = "wwz")
-w2000 <- decomp(wid2000,
-                wfd2000,
+w2000 <- decomp(inter2000,
+                final2000,
                 countries,
-                sectors,
+                industries,
                 output2000,
                 method = "wwz")
-w2005 <- decomp(wid2005,
-                wfd2005,
+w2005 <- decomp(inter2005,
+                final2005,
                 countries,
-                sectors,
+                industries,
                 output2005,
                 method = "wwz")
-w2008 <- decomp(wid2008,
-                wfd2008,
+w2008 <- decomp(inter2008,
+                final2008,
                 countries,
-                sectors,
[…]

### 2016-01-27 Draft.tex

\documentclass[11pt,a4paper]{article}

\usepackage{mathtools}
\usepackage{threeparttable}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{changepage}
\usepackage{verbatim}
\usepackage{epigraph}
\usepackage{verbatim}
\usepackage{lscape}
\usepackage{pbox}
\usepackage{array}
\usepackage{natbib}

\linespread{1.5}

\begin{document}

\section{Introduction}
The emergence of Global Value Chains (GVCs) offers a new path to industrialisation for developing countries. As \cite{riba12} phrases it, internationally fragmented production allows developing countries to join existing supply chains instead of building them. This brings about many potential advantages for these countries. Connecting with firms from advanced nations allows developing nations, for instance, to benefit from their sophisticated technologies and know-how. In addition, relying on an existing production network frees them from constraints imposed by economies of scale and the increased specialisation that GVCs imply limits the negative impact of unproductive parts of the domestic supply chain. After all, when competition moves from goods to tasks, comparative advantage becomes much finer and does not require a broad range of productive stages domestically. Conditional evidence for such a positive impact of GVC participation in low- and middle-income countries is presented in \citet{viku16} and \citet{unct13}.

Empirically, the considerable expansion of GVCs has been documented in several recent studies. For instance, \citet{dahuetal98, dahuetal01} show in two early seminal contributions that GVCs are responsible for a major share of the total growth in world trade from 1970 to 1990. Amongst others, \citet{rojoguno12a} and \citet{ribajalo13} find that this growth in GVC trade has even accelerated in the recent two decades. Furthermore, this work has not only revealed a rapid rise in production fragmentation across borders but it has also re-evaluated important indicators of trade, such as bilateral trade imbalances and revealed comparative advantage showing that calculating GVC indicators is central to a better understanding of countries' trade patters and competitiveness.

A central step towards a more in-depth analysis of GVCs has been laid by \citet{rokoetal14} and \citet{zhwaetal13} who show that it is necessary to go beyond deriving origins of value added to examine production sharing comprehensively. They split exports into different categories and calculate metrics of how often goods cross borders. This enables them to derive measures of GVC length but also allows to investigate how individual countries are integrated into GVCs. For instance, they show that a considerable part of US value added exports eventually returns home in the form of final goods which is indicative of the US being specialised in upstream production stages.\footnote{See \citeauthor{joamsoca15} (forthcoming) for a comprehensive review of the literature on GVCs and outsourcing.}

However, these contributions typically have two shortcomings. Firstly, most evidence is based on data from high-income countries. The reason is that reliable time-series of both national and international input-output tables have only been available for this particular subset of countries. In addition, the evidence is regularly based on a small sample of GVC indicators that hide valuable information stemming from more decomposed and disaggregated indicators.

In this paper we address these issues by applying the novel and more detailed gross export decomposition developed by \citet{zhwaetal13} and \citet{rokoetal14} to a new set of Inter-Country Input-Output tables (ICIOs) with extensive country coverage provided by the OECD. The new ICIOs allow us to get a better understanding of the GVC activities of low- and middle-income countries while the new decomposition allows us to zoom in more closely at these activities revealing information not available from standard GVC indicators.

Our analysis confirms the expansion of GVCs in recent years and presents evidence that GVCs have become longer over time. However, we also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level.

The paper is organised as follows. Section \ref{sec:data} reviews shortly the decomposition proposed by \citet[WWZ henceforth]{zhwaetal13} and outlines the new ICIOs provided by the OECD. Section \ref{sec:basics} discusses standard indicators and measures calculated with the new data while section \ref{sec:news} discusses the results for the novel indicators. Section \ref{sec:conclusion} concludes.

\section[New data and new indicators]{New data and new indicators\footnote{The following section draws heavily from \citet{zhwaetal13}, \citet{viku16}, and \cite{baquviku15}.}}\label{sec:data}

GVC analysis relies typically on Inter-Country Input-Output tables (ICIOs). ICIOs are matrices that give supply and demand relationships between industries within and across countries. For instance, ICIOs state the amount of inputs of the Indian steel industry in the output of the US car industry. However, for a correct examination of GVCs it is necessary to go a step further from the ICIOs by deriving the true value added origins of the US car output. If, for example, India depends on inputs from the US steel industry to supply the US car industry, then ICIOs overstate the actual contribution of India. The extension of the basic \citet{wale36} insight by \citet{dahuetal01} shows how the information in ICIOs can be decomposed to estimate such value added flows.

The idea is that the production of industry \textit{i} creates value added in industry $i$ itself, a direct contribution, but also in industries that supply $i$ with inputs, an indirect contribution. Since these industries themselves rely on inputs, $i$'s production sets several rounds of indirect value added creation in motion that can mathematically be expressed as:
\begin{equation}
VB = V + VA + VAA + VAAA + ... = V (I+A+A^{2}+A^{3}+...),
\end{equation}
which, as an infinite geometric series with the elements of $A<1$, simplifies to
\begin{equation}
VB = V (I-A)^{-1},
\end{equation}
where \textit{V} is a matrix with the diagonal representing the direct value added contribution of each industry, \textit{A} is the Input-Output coefficient matrix, which means it gives the direct input flows between industries required for 1\$ of output, and $B = (I-A)^{-1}$ is the so called Leontief inverse. \textit{VB} gives thus so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries. If we post-multiply $VB$ with exports, we get a matrix with the value added origins of each industry's exports.

This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures, namely foreign value added in exports (Vertical Specialisation) and domestic value added in foreign exports (Vertical Specialisation 1). These indicators can tell us how much a country is integrated in GVCs and if it acts mainly as a supplier or a host of value added. However, the Leontief decomposition is only informative for the origin and destination of value added while ICIOs also contain info on the type of good that is being traded and and how often an intermediate crosses borders. The WWZ decomposition extends the Leontief decomposition in this direction and thereby extracts more insights from ICIOs.

\subsection{Wang-Wei-Zhu decomposition}\label{sub:wwz}

Since the derivation itself is not the focus of this paper, we present here only the final result for a $G$-country $N$-industry model (equation 37 in WWZ) and refer the interested reader to the original paper. WWZ use the Leontief decomposition and extend it using additional information from ICIOs on the final usage and destination of the exports (e.g. re-imported vs. absorbed abroad). This splits the exports, $E$, of industry $l$ in country $k$ into sixteen different parts broadly differentiated into the four boroad categories domestic value added absorbed abroad, domestic value added returning home, foreign value added, and purely double counted terms:
\begin{align}\label{eq:wwz}
\begin{split}
E^{kl}
= &\left(V^k B^{kk} \right)^T * F^{kl} 
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{ll} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tt} )
+ \left(V^k L^{kk} \right)^T *  (A^{kl} B^{ll} \sum_{t \neq k,l}^G  F^{lt} )\\ 
+&  \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G \sum_{l,u \neq k,t}^G B^{lt} F^{tu} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{ll} F^{lk} \right)\\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k,l}^G  B^{lt} F^{tk} )
+ \left(V^k L^{kk} \right)^T * \left(A^{kl} B^{lk} F^{kk} \right) \\
+& \left(V^k L^{kk} \right)^T * (A^{kl} \sum_{t \neq k}^G  B^{lk} F^{kt} )
+ \left(V^k B^{kk} -  V^k L^{kk} \right)^T * \left(A^{kl} X^{l}  \right)\\
+& \left(V^l B^{lk} \right)^T * F^{kl}
+ \left(V^l B^{lk} \right)^T *  \left(A^{kl} L^{ll} F^{ll} \right)
+ \left(V^l B^{lk} \right)^T \\
*&  \left(A^{kl} L^{ll} E^{l*} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} * F^{kl}
+ (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T}\\
*&   \left(A^{kl} L^{ll} F^{ll} \right) + (\sum_{t \neq k,l}^G  V^{t} B^{tk} )^{T} *  \left(A^{kl} L^{ll} E^{l*} \right) ,
\end{split}
\end{align}
where $F$ is final demand, and $L$ refers to the domestic Leontief inverse as opposed to the global inverse $B$. $X$ is output while \textit{T} indicates a matrix transpose operation.

The four main categories are further divided according to their final destination so that the final decomposition is given by:

\begin{itemize}
\item Domestic value added absorbed abroad (\textit{VAX\_G}, T1-5)
\begin{itemize}
\item Domestic value added in final exports (\textit{DVA\_FIN}, T1)
\item Domestic value added in intermediate exports
\begin{itemize}
\item Domestic value added in intermediate exports absorbed by direct importers (\textit{DVA\_INT}, T2)
\item Domestic value added in intermediate exports re-exported to third countries (\textit{DVA\_INTrex}, T3-5)
\begin{itemize}
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce domestic final goods (\textit{DVA\_INTrexI1}, T3)
\item Domestic value added in intermediate exports re-exported to third countries as  final goods (\textit{DVA\_INTrexF}, T4)
\item Domestic value added in intermediate exports re-exported to third countries as intermediate goods to produce exports (\textit{DVA\_INTrexI2}, T5)
\end{itemize}
\end{itemize}
\end{itemize}
\item Domestic value added returning home (\textit{RDV\_B}, T6-8)
\begin{itemize}
\item Domestic value added returning home as final goods (\textit{RDV\_FIN}, T6)
\item Domestic value added returning home as final goods through third countries (\textit{RDV\_FIN2}, T7)
\item Domestic value added returning home as intermediate goods (\textit{RDV\_INT}, T8)
\end{itemize}
\item Foreign value added (\textit{FVA}, T11-12/14-15 )
\begin{itemize}
\item Foreign value added in final good exports (\textit{FVA\_FIN}, T11/14)
\begin{itemize}
\item Foreign value added in final good exports sourced from direct importer (\textit{MVA\_FIN}, T11)
\item Foreign value added in final good exports sourced from other countries (\textit{OVA\_FIN}, T14)
\end{itemize}
\item Foreign value added in intermediate good exports (\textit{FVA\_INT}, T12/15)
\begin{itemize}
\item Foreign value added in intermediate good exports sourced from direct importer (\textit{MVA\_INT}, T12)
\item Foreign value added in intermediate good exports sourced from other countries(\textit{OVA\_INT}, T15)
\end{itemize}
\end{itemize}
\item Pure double counting (\textit{PDC}, T9-10/13/16)
\begin{itemize}
\item Pure double counting from domestic source (\textit{DDC}, T9-10)
\begin{itemize}
\item Due to final goods exports production (\textit{DDF}, T9)
\item Due to intermediate goods exports production (\textit{DDI}, T10)
\end{itemize}
\item Pure double counting from foreign source (\textit{FDC}, T13/16)
\begin{itemize}
\item Due to direct importer exports production (\textit{FDF}, T13)
\item Due to other countries' exports production (\textit{FDI}, T16)
\end{itemize}
\end{itemize}
\end{itemize}

The advantage of such a detailed decomposition is that these new indicators can inform us on how countries integrate into GVCs while the basic Leontief decomposition mainly informs us on the intensity of integration. High amounts of foreign value added in final goods exports are, for instance, suggestive of a specialisation in downstream tasks that add little value to a good, such as assembly. High amounts of domestic value added in intermediate exports, on the other hand, are evidence of a more upstream specialisation in tasks that aaa a lot of value, such as business services. By tracking these two variables over time we can then see which countries have succeeded in moving up the value chain.

The high resolution of the WWZ decomposition does, however, not mean that the Leontief decomposition does not contain valuable information at all. In fact, we exploit the decomposition of exports into source industry and source country by calculating variants of the standard indicators based on different characteristics. In particular, we will assess the integration of low- and middle-income countries into GVCs by computing the amount of value added that they supply for total GVC trade.

\subsection{OECD ICIOs}\label{sub:icios}
We use the new OECD ICIOs as the main data source for the GVC indicators and the industry position indicators. The OECD ICIOs constitute the most recent and most advanced release of Inter-Country Input-Output tables. The new version of the database provides ICIOs covering 61 countries and 34 industries for the years 1995, 2000, 2005, and 2008 to 2011.\footnote{Countries and industries are listed in the Appendix.}\textsuperscript{,}\footnote{Note that in the analysis 2009 and 2010 are excluded due to the global crisis.} This extensive country coverage is crucial for analysing how GVCs affect countries at different stages of development over time, a feature that has not been possible due to limited data availability in previous databases. The empirical literature discussed above shows that especially the extended coverage of Asia is important. To create ICIOs, the OECD combines national IO tables with international trade data. As OECD countries have a harmonised construction methodology, potential discrepancies between national IO tables should be minor. Furthermore, the advanced harmonisation across countries reduces the use of proportionality assumptions to derive the ratio of imported intermediates in an industry's demand to a minimum. In addition, the OECD has used elabourate techniques to deal with China's processing trade. Due to China's outstanding role in GVCs and processing trade, this implies a significant improvement for the reliability of the database.\footnote{See \citet{rokoetal12} for an analysis of China's processing trade.}

\section{What we know: Old facts with new data}\label{sec:basics}

\section{New trends and patterns in GVCs}\label{sec:news}

\section{Conclusion}\label{sec:conclusion}

\clearpage

\bibliographystyle{chicago}
\bibliography{GVCandLMICs}

\clearpage

\end{document}

### 2016-01-27 R/decompose.R
add instructions for wwz, organise file

@@ -5,7 +5,10 @@
 # load the library
 library(decompr)
 
-# run the Leontief decomposition
+
+# run the decompositions and save the output
+
+## run the Leontief decomposition
 l1995 <- decomp(inter1995,
                 final1995,
                 countries,
@@ -49,12 +52,17 @@ l2011 <- decomp(inter2011,
                 output2011,
                 method = "leontief")
 
-# save Leontief the data
-save(l1995, l2000, l2005, l2008, file = "data/TiVa-Leontief.RData")
+## save Leontief the data
+save(l1995,
+     l2000,
+     l2005,
+     l2008,
+     l2009,
+     l2010,
+     l2011,
+     file = "data/TiVa-Leontief.RData")
 
-
-
-# run the Wang-Wei-Zhu decomposition
+## run the Wang-Wei-Zhu decomposition
 w1995 <- decomp(inter1995,
                 final1995,
                 countries,
@@ -79,6 +87,31 @@ w2008 <- decomp(inter2008,
                 industries,
                 output2008,
                 method = "wwz")
+w2009 <- decomp(inter2009,
+                final2009,
+                countries,
+                industries,
+                output2009,
+                method = "wwz")
+w2010 <- decomp(inter2010,
+                final2010,
+                countries,
+                industries,
+                output2010,
+                method = "wwz")
+w2011 <- decomp(inter2011,
+                final2011,
+                countries,
+                industries,
+                output2011,
+                method = "wwz")
 
 # save the Wang-Wei-Zhu data
-save(w1995, w2000, w2005, w2008, file = "data/TiVa-WWZ.RData")
+save(w1995,
+     w2000,
+     w2005,
+     w2008,
+     w2009,
+     w2010,
+     w2011,
+     file = "data/TiVa-WWZ.RData")

### 2016-01-28 R/merge-wwz.R
update to 2011

@@ -5,31 +5,44 @@
 # load libraries
 library(magrittr)
 
+
 # load the data
 load("data/w1995.RData")
 load("data/w2000.RData")
 load("data/w2005.RData")
 load("data/w2008.RData")
+load("data/w2009.RData")
+load("data/w2010.RData")
+load("data/w2011.RData")
+
 
 # add year variable to each data set
 w1995 <- cbind(year = 1995, w1995)
 w2000 <- cbind(year = 2000, w2000)
 w2005 <- cbind(year = 2005, w2005)
 w2008 <- cbind(year = 2008, w2008)
+w2009 <- cbind(year = 2009, w2009)
+w2010 <- cbind(year = 2010, w2010)
+w2011 <- cbind(year = 2011, w2011)
+
 
 # stack the data sets
-w1995_2008 <- rbind(w1995, w2000, w2005, w2008)
+w1995_2011 <- rbind(w1995, w2000, w2005, w2008, w2009, w2010, w2011)
 
-# # to lower case
-levels(w1995_2008$Exporting_Country) <- tolower(levels(w1995_2008$Exporting_Country))
-levels(w1995_2008$Exporting_Industry) <- tolower(levels(w1995_2008$Exporting_Industry))
-levels(w1995_2008$Importing_Country) <- tolower(levels(w1995_2008$Importing_Country))
+
+# to lower case
+levels(w1995_2011$Exporting_Country) <- tolower(levels(w1995_2011$Exporting_Country))
+levels(w1995_2011$Exporting_Industry) <- tolower(levels(w1995_2011$Exporting_Industry))
+levels(w1995_2011$Importing_Country) <- tolower(levels(w1995_2011$Importing_Country))
+
 
 # remove ROW (Rest of World)
-w1995_2008 %<>% subset(Exporting_Country != "row")
+w1995_2011 %<>% subset(Exporting_Country != "row")
+
 
 # remove unneeded columns
-w1995_2008 %<>% select(-(24:29))
+w1995_2011 %<>% select(-(24:29))
+
 
 # save
-save(w1995_2008, file = "data/w1995_2008.RData")
+save(w1995_2011, file = "data/w1995_2011.RData")

### 2016-01-28 R/analyse_gvc_indicators.R
update to 2011

@@ -6,7 +6,7 @@
 load("data/gvc_indicators.RData")
 load("data/nrca_df.RData")
 load("data/country_vars.RData")
-load("data/w1995_2008.RData")
+load("data/w1995_2011.RData")
 load("data/trca.RData")
 
 # load required libraries
@@ -27,7 +27,7 @@ country_vars$ic      <- with(country_vars, ifelse(avg_gdppc <= 6000, "lic",
                                                   ifelse(avg_gdppc > 12000, "hic", "mic") ) )
 
 # merge wwz and country vars
-w1995_2008 %<>% merge(country_vars,
+w1995_2011 %<>% merge(country_vars,
                       by.x = c("year", "Exporting_Country"),
                       by.y = c("year", "ctry")              )
 
@@ -186,21 +186,21 @@ gvc_indicators %>%
 
 # plot gvc_length
 # RDV
-w1995_2008 %>%
+w1995_2011 %>%
   group_by(year) %>%
   summarise( RDV = () sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) / sum(texp) ) %>%
   ggvis(~year, ~RDV) %>%
   layer_lines()
 
 # PDC
-w1995_2008 %>%
+w1995_2011 %>%
   group_by(year) %>%
   summarise( PDC = ( sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) / sum(texp) ) %>%
   ggvis(~year, ~PDC) %>%
   layer_lines()
 
 # both
-w1995_2008 %>%
+w1995_2011 %>%
   group_by(year) %>%
   summarise( RDV = ( sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) / sum(texp),
              PDC = ( sum(DDC_FIN) + sum(DDC_INT) + sum(ODC) + sum(MDC) ) / sum(texp) ) %>%
@@ -291,14 +291,14 @@ gvc_indicators %>%
 ## WWZ by IC
 
 # RDV
-w1995_2008 %>%
+w1995_2011 %>%
   group_by(year, ic) %>%
   summarise( RDV = ( sum(RDV_INT) + sum(RDV_FIN) + sum(RDV_FIN2) ) / sum(texp) ) %>%
   ggvis(~year, ~RDV, stroke = ~ic) %>%
[…]

### 2016-04-07 R/analyse_gvc_indicators.R
add graph with divided by exports

@@ -200,6 +200,21 @@ gvc_indicators %>%
   ggvis(~year, ~i2e ) %>%
   layer_lines()
 
+# i2e with divided
+# i2e with divided
+gvc_indicators %>%
+  group_by(year) %>%
+  summarise(i2e = sum(i2e) ) %>%
+  mutate(divided = c(17.3, 20.5, 22.1, 23.2, 21.0, 22.4, 23.2)) %>%
+  ggvis(x=~year) %>%
+  layer_lines(y=~i2e) %>%
+  add_axis('y', orient = 'left', title = 'i2e') %>%
+  add_axis('y', 'ydiv' , orient = 'right',
+           title= 'i2e as Percentage of exports (red)', grid=F, title_offset = 50,
+           properties = axis_props(labels = list(fill = 'red')) ) %>%
+  layer_lines( prop('y', ~divided, scale='ydiv'), stroke:='red' )
+
+
 # i2e class == L
 gvc_indicators %>%
   group_by(year) %>%

### 2016-04-07 Draft.tex

@@
 Our analysis confirms the expansion of GVCs in recent years and presents evidence that GVCs have become longer over time.
-However, we also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level.
+We also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level.
 
@@
 
+In this section we confirm standard stylized facts of GVC integration that are typically based on smaller samples with OECD ICIO data.
+We starting by look at the development of two basic measures of GVC integration introduced by \citet{dahuetal01}, namely $fvax$ and $dvar$.
+Both show the well-known increase
+
 \section{New trends and patterns in GVCs}\label{sec:news}

### 2016-04-08 Draft.tex
Draft update
Chapter 3 60% done

@@
 \usepackage{epigraph}
-\usepackage{verbatim}
 \usepackage{lscape}
@@
 \usepackage{natbib}
+\usepackage{graphicx}
 
 \linespread{1.5}
+\graphicspath{{/Users/VK/Documents/Github/OECD-TiVA-LMIC-GVCs/misc/}}
 
@@
 Furthermore, the advanced harmonisation across countries reduces the use of proportionality assumptions to derive the ratio of imported intermediates in an industry's demand to a minimum.
-In addition, the OECD has used elabourate techniques to deal with China's processing trade.
+In addition, the OECD has used elaborate techniques to deal with China's processing trade.
 Due to China's outstanding role in GVCs and processing trade, this implies a significant improvement for the reliability of the database.\footnote{See \citet{rokoetal12} for an analysis of China's processing trade.}
@@
 
-In this section we confirm standard stylized facts of GVC integration that are typically based on smaller samples with OECD ICIO data.
-We starting by look at the development of two basic measures of GVC integration introduced by \citet{dahuetal01}, namely $fvax$ and $dvar$.
-Both show the well-known increase
+In this section we confirm with OECD ICIO data standard stylised facts of GVC integration that are typically based on smaller samples.
+We starting by look at the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports or $fvax$, which captures backward linkages into value chains.
+It shows the well-known increase in GVC integration from 1995 to 2011.
+As illustrated in Figure \ref{fig:i2e_t}, the value of $fvax$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%.
+Thus, countries rely for their export production increasingly on inputs produced abroad.
+After the slump during the financial crises in 2009, GVCs seem to have quickly recovered and already have started to exceed their pre-crisis levels in 2011.
+
+\begin{figure}
+\centering
+\label{fig:i2e_t}
+\includegraphics[scale=0.4]{i2epercentage.png}
+\caption{The development of GVC integration over time}
+\end{figure}
+
+Looking at sectoral $fvax$ averages, Figure \ref{fig:i2e_s} reveals that the sectors exhibiting the highest dependency on foreign inputs are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ).
+In particular, the transport equipment and electronics industry are strongly engaged in GVCs having highly international production networks.
+For instance, Apple's iPhone contains inputs from 9 to 10 countries while the Boeing 787 production spans more than 5 countries.
+These sectors are typically close to final demand or depend on imported raw materials.
+
+The bottom 6 industries are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT).
+These sectors are typically located upstream in the supply chain far from final demand and have high high value added to output ratios.
+
+Naturally then, things are reversed when we look at the corresponding forward linkage GVC measure, $dvar$.
+It captures the amount of domestic value added in foreign exports and thus quantifies how important domestic industries are for foreign export production.
+Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $fvax$ ranking such as mining or financial and telecommunication services (FIN and PTL).
+This shows that these industries are also strongly engaged in GVCs but their participation is of a different type.
+They primarily supply important inputs but do not serve final demand.
+
+With respect to services, it is also indicative of the servicification of manufacturing as described by \citet{ribaetal15}.
+This means that an increasing share of manufacturing gross exports is actually value added generated in services sectors and then embedded in the intermediate exports of manufacturers.
+This importance of services sectors for exports cannot be seen from standard gross trade statistics and thus constitutes a major advantage of trade in value added measures.
+It is also indicative of a growing internationalisation of services.
+Increasingly, services are being offshored and sourced from abroad.
+In that respect, it is also interesting to note that despite the low absolute $fvax$ shares, it is in services where a lot of the growth in $fvax$ has taken place.
+As can be seen in Figure \ref{fig:i2e_st}, five out of the six sectors with the highest growth in $fvax$ shares are services sectors.
 

### 2016-05-19 R/last_minute_stuff.R
add plots

@@ -8,6 +8,7 @@ library(dplyr)
 
 # load data
 sector_desc <- read_excel(path = 'data/sector_desc.xlsx')
+wwz         <- read_excel(path = 'data/graphs.xlsx')
 
 # plot top 6
 sector_desc %>%
@@ -28,3 +29,21 @@ sector_desc %>%
   ggvis(x=~sector, y=~fvax) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+wwz1 <- wwz[2:8,1:3]
+names(wwz1) <- c('year', 'e2rL-LM', 'i2eL-LM')
+wwz1$year <- as.numeric(wwz1$year)
+wwz2 <- wwz[11:17,]
+names(wwz2) <- wwz[10,]
+names(wwz2)[1] <- 'year'
+wwz2$year      <- as.numeric(wwz2$year)
+wwz2$fva_fin   <- as.numeric(wwz2$fva_fin)
+wwz2$fva_inter <- as.numeric(wwz2$fva_inter)
+wwz2$dva_fin   <- as.numeric(wwz2$dva_fin)
+wwz2$dva_inter <- as.numeric(wwz2$dva_inter)
+wwz2$rdv       <- as.numeric(wwz2$rdv)
+
+wwz1 %>%
+  ggvis(x=~year, y=~`e2rL-LM`, stroke='e2r') %>%
+  layer_lines() %>%
+  layer_lines(x=~year, y=~`i2eL-LM`, stroke='i2e')

### 2016-05-19 R/last_minute_stuff.R
add plots

@@ -33,6 +33,10 @@ sector_desc %>%
 wwz1 <- wwz[2:8,1:3]
 names(wwz1) <- c('year', 'e2rL-LM', 'i2eL-LM')
 wwz1$year <- as.numeric(wwz1$year)
+wwz1$`e2rL-LM` <- as.numeric(wwz1$`e2rL-LM`)
+wwz1$`i2eL-LM` <- as.numeric(wwz1$`i2eL-LM`)
+wwz1$`e2rL-LM` <- round(wwz1$`e2rL-LM`)
+
 wwz2 <- wwz[11:17,]
 names(wwz2) <- wwz[10,]
 names(wwz2)[1] <- 'year'
@@ -47,3 +51,10 @@ wwz1 %>%
   ggvis(x=~year, y=~`e2rL-LM`, stroke='e2r') %>%
   layer_lines() %>%
   layer_lines(x=~year, y=~`i2eL-LM`, stroke='i2e')
+
+wwz2 %>%
+  ggvis(x=~year, y=~fva_fin, stroke='fva_fin') %>%
+  layer_lines() %>%
+  layer_lines(x=~year, y=~fva_inter, stroke='fva_inter') %>%
+  layer_lines(x=~year, y=~dva_fin,   stroke='dva_fin')   %>%
+  layer_lines(x=~year, y=~dva_inter, stroke='dva_inter')

### 2016-05-19 R/last_minute_stuff.R
add pdc plot

@@ -9,6 +9,7 @@ library(dplyr)
 # load data
 sector_desc <- read_excel(path = 'data/sector_desc.xlsx')
 wwz         <- read_excel(path = 'data/graphs.xlsx')
+pdc         <- read_excel(path = 'data/graphs2.xlsx')
 
 # plot top 6
 sector_desc %>%
@@ -58,3 +59,8 @@ wwz2 %>%
   layer_lines(x=~year, y=~fva_inter, stroke='fva_inter') %>%
   layer_lines(x=~year, y=~dva_fin,   stroke='dva_fin')   %>%
   layer_lines(x=~year, y=~dva_inter, stroke='dva_inter')
+
+pdc %>%
+  ggvis(x=~year, y=~pdc_t, stroke='pdc_t') %>%
+  layer_lines() %>%
+  layer_lines(x=~year, y=~pdc_exp, stroke='pdc_exp')

### 2016-05-19 R/last_minute_stuff.R
pdf use differentiated y axis

@@ -61,6 +61,10 @@ wwz2 %>%
   layer_lines(x=~year, y=~dva_inter, stroke='dva_inter')
 
 pdc %>%
-  ggvis(x=~year, y=~pdc_t, stroke='pdc_t') %>%
-  layer_lines() %>%
-  layer_lines(x=~year, y=~pdc_exp, stroke='pdc_exp')
+  ggvis(x=~year) %>%
+  layer_lines(y=~pdc_t) %>%
+  add_axis('y', orient = 'left', title = 'pdc_t') %>%
+  add_axis('y', 'ydiv' , orient = 'right',
+           title= 'i2e as Percentage of exports (red)', grid=F, title_offset = 50,
+           properties = axis_props(labels = list(fill = 'red')) ) %>%
+  layer_lines( prop('y', ~pdc_exp, scale='ydiv'), stroke:='red' )

### 2016-05-19 R/last_minute_stuff.R
add plots and rename

@@ -14,20 +14,30 @@ pdc         <- read_excel(path = 'data/graphs2.xlsx')
 # plot top 6
 sector_desc %>%
   group_by(sector) %>%
-  summarise(fvax=fvax_s/exp) %>%
-  arrange(fvax) %>%
-  filter(fvax > 0.30) %>%
-  ggvis(x=~sector, y=~fvax) %>%
+  summarise(i2e=fvax_s/exp) %>%
+  arrange(i2e) %>%
+  filter(i2e > 0.30) %>%
+  ggvis(x=~sector, y=~i2e) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
 
 # plot bottom 6
 sector_desc %>%
   group_by(sector) %>%
-  summarise(fvax=fvax_s/exp) %>%
-  arrange(fvax) %>%
-  filter(fvax < 0.13) %>%
-  ggvis(x=~sector, y=~fvax) %>%
+  summarise(i2e=fvax_s/exp) %>%
+  arrange(i2e) %>%
+  filter(i2e < 0.13) %>%
+  ggvis(x=~sector, y=~i2e) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# plot top 6 and bottom 6
+sector_desc %>%
+  group_by(sector) %>%
+  summarise(i2e=fvax_s/exp) %>%
+  arrange(i2e) %>%
+  filter(i2e > 0.30 | i2e < 0.13) %>%
+  ggvis(x=~sector, y=~i2e) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
 
@@ -65,6 +75,6 @@ pdc %>%
   layer_lines(y=~pdc_t) %>%
   add_axis('y', orient = 'left', title = 'pdc_t') %>%
   add_axis('y', 'ydiv' , orient = 'right',
-           title= 'i2e as Percentage of exports (red)', grid=F, title_offset = 50,
+           title= 'pdc_exp', grid=F, title_offset = 50,
            properties = axis_props(labels = list(fill = 'red')) ) %>%
   layer_lines( prop('y', ~pdc_exp, scale='ydiv'), stroke:='red' )

### 2016-05-19 R/last_minute_stuff.R
better titles

@@ -11,7 +11,7 @@ sector_desc <- read_excel(path = 'data/sector_desc.xlsx')
 wwz         <- read_excel(path = 'data/graphs.xlsx')
 pdc         <- read_excel(path = 'data/graphs2.xlsx')
 
-# plot top 6
+# i2e plot top 6
 sector_desc %>%
   group_by(sector) %>%
   summarise(i2e=fvax_s/exp) %>%
@@ -21,7 +21,7 @@ sector_desc %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
 
-# plot bottom 6
+# i2e plot bottom 6
 sector_desc %>%
   group_by(sector) %>%
   summarise(i2e=fvax_s/exp) %>%
@@ -31,13 +31,43 @@ sector_desc %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
 
-# plot top 6 and bottom 6
+# i2e plot top 6 and bottom 6
 sector_desc %>%
   group_by(sector) %>%
   summarise(i2e=fvax_s/exp) %>%
   arrange(i2e) %>%
   filter(i2e > 0.30 | i2e < 0.13) %>%
   ggvis(x=~sector, y=~i2e) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# e2r plot top 6
+sector_desc %>%
+  group_by(sector) %>%
+  summarise(e2r=fvax_s/exp) %>%
+  arrange(e2r) %>%
+  filter(e2r > 0.30) %>%
+  ggvis(x=~sector, y=~e2r) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# e2r plot bottom 6
+sector_desc %>%
+  group_by(sector) %>%
+  summarise(e2r=fvax_s/exp) %>%
+  arrange(e2r) %>%
+  filter(e2r < 0.13) %>%
+  ggvis(x=~sector, y=~e2r) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# e2r plot top 6 and bottom 6
+sector_desc %>%
+  group_by(sector) %>%
+  summarise(e2r=fvax_s/exp) %>%
+  arrange(e2r) %>%
[…]

### 2016-05-19 Draft.tex
updated draft

@@
 \usepackage{longtable}
+\usepackage{multirow}
 \usepackage{booktabs}
@@
 We also find that these developments are increasingly driven by low- and middle-income countries while the integration of high-income countries has begun to even out at a high level.
+In addition, we find that high-income countries typically are the starting and end points of GVCs in that they provide upstream inputs and then serve eventually again as demand markets for the final products.
+Low- and middle-income countries, on the other hand, are rather specialised in downstream activities such as assembly and export typically less domestic value added.
+However, we also observe that developing economies move more and more upstream along the value chain and should thus be able to generate more gains from their GVC participation.
 
@@
 Section \ref{sec:data} reviews shortly the decomposition proposed by \citet[WWZ henceforth]{zhwaetal13} and outlines the new ICIOs provided by the OECD.
-Section \ref{sec:basics} discusses standard indicators and measures calculated with the new data while section \ref{sec:news} discusses the results for the novel indicators.
+Section \ref{sec:basics} discusses results using standard indicators and measures calculated with the new data while section \ref{sec:news} discusses the results for the novel indicators.
 Section \ref{sec:conclusion} concludes.
@@
 
-This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures, namely foreign value added in exports (Vertical Specialisation) and domestic value added in foreign exports (Vertical Specialisation 1).
-These indicators can tell us how much a country is integrated in GVCs and if it acts mainly as a supplier or a host of value added.
+This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures, namely foreign value added in exports (Hummels et al.'s Vertical Specialisation) and domestic value added in foreign exports (Hummels et al.'s Vertical Specialisation 1).
+These indicators can tell us how much a country is integrated into GVCs and if it acts mainly as a supplier or a user of foreign value added.
 However, the Leontief decomposition is only informative for the origin and destination of value added while ICIOs also contain info on the type of good that is being traded and and how often an intermediate crosses borders.
@@
 
-In this section we confirm with OECD ICIO data standard stylised facts of GVC integration that are typically based on smaller samples.
-We starting by look at the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports or $fvax$, which captures backward linkages into value chains.
-It shows the well-known increase in GVC integration from 1995 to 2011.
-As illustrated in Figure \ref{fig:i2e_t}, the value of $fvax$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%.
+In this section we look into stylised facts on GVC integration that are typically based on smaller samples with our extensive OECD ICIO dataset.
+We start by examining the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward and backward linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011.
+As illustrated in Figure \ref{fig:i2e_t}, the value of $i2e$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%.
 Thus, countries rely for their export production increasingly on inputs produced abroad.
-After the slump during the financial crises in 2009, GVCs seem to have quickly recovered and already have started to exceed their pre-crisis levels in 2011.
+The numbers are in line with similar findings by \citet{rojoguno12b} but their sample ends in 2009.
+It is then interesting to see that after the slump during the financial crises in 2009, GVCs have quickly recovered and already have started to exceed their pre-crisis levels in 2011.
 
@@
 
-Looking at sectoral $fvax$ averages, Figure \ref{fig:i2e_s} reveals that the sectors exhibiting the highest dependency on foreign inputs are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ).
+Another way to examine the expansion of GVCs from 1995 to 2011 is to look at their intensity or length instead of their trade volume.
+\citet{zhwaetal13} propose to use the amount of double counted trade, $pdc$, as a proxy for GVC length since its value goes up the more back-and-forth trade occurs, which is equivalent to an increase in the number of production stages.
+They show that its value has increased for the 40 WIOD countries.
+In Figure \ref{fig:pdc_t}, we observe similarly that $pdc$ as a share of total exports has increased over the examined period by 73\% and thus more than $i2e$.
+Therefore, GVCs do not only channel more trade but also have become longer over time.
+
+%pdc change over time figure
+
+Turning from the development over time to sectoral differences in GVC integration, Figure \ref{fig:i2e_s} reveals in line with \citet{rojoguno12a} that the sectors exhibiting the highest dependency on foreign inputs in terms of $i2e$ shares are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ).
 In particular, the transport equipment and electronics industry are strongly engaged in GVCs having highly international production networks.
 For instance, Apple's iPhone contains inputs from 9 to 10 countries while the Boeing 787 production spans more than 5 countries.
-These sectors are typically close to final demand or depend on imported raw materials.
+The sectors can be characterised as being close to final demand or being dependent on imported raw materials.
+These characteristics can thus explain differences in GVC integration.
 
-The bottom 6 industries are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT).
+The bottom 6 industries in terms of $i2e$ shares are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT).
 These sectors are typically located upstream in the supply chain far from final demand and have high high value added to output ratios.
 
-Naturally then, things are reversed when we look at the corresponding forward linkage GVC measure, $dvar$.
+% sectoral graph i2e
+
+Naturally then, things are reversed when we look at the corresponding forward linkage GVC measure, $e2r$.
 It captures the amount of domestic value added in foreign exports and thus quantifies how important domestic industries are for foreign export production.
-Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $fvax$ ranking such as mining or financial and telecommunication services (FIN and PTL).
+Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $i2e$ ranking such as mining or financial and telecommunication services (FIN and PTL).
 This shows that these industries are also strongly engaged in GVCs but their participation is of a different type.
@@
 Increasingly, services are being offshored and sourced from abroad.
-In that respect, it is also interesting to note that despite the low absolute $fvax$ shares, it is in services where a lot of the growth in $fvax$ has taken place.
-As can be seen in Figure \ref{fig:i2e_st}, five out of the six sectors with the highest growth in $fvax$ shares are services sectors.
+In that respect, it is also interesting to note that despite the low absolute $i2e$ shares, it is in services where a lot of the growth in $i2e$ has taken place.
+As can be seen in Figure \ref{fig:i2e_st}, five out of the six sectors with the highest growth in $i2e$ shares are services sectors.
 
-\section{New trends and patterns in GVCs}\label{sec:news}
+% sectoral graph e2r
+
+Finally, when we turn to differences in GVC integration by country we can confirm the findings by \citet{ribajalo13}.
+Figure \ref{i2e_k} shows that small countries close to the major GVC hubs in Asia, Europe, and North America such as Slovakia or Malaysia have the highest average $i2e$ shares.
+Countries specialised in the primary sector or assembly on the other hand have very low values.
+Correspondingly, Latin American countries with their focus on agriculture and mining have very weak backward linkages into GVCs.
+However, the development over time shows that some of the countries with the relatively low GVC integration have begun to catch up.
+For instance, Argentina, Indian and Turkey are in the top 6 when it comes to the growth of $i2e$ shares from 1995 to 2011.
+
+%country graph i2e
+
+Driven by the sectoral statistics, we then find again that for $e2r$ the picture is reversed with raw material exporters on top.
+If we abstract from these countries we find technologically advanced countries such as Switzerland and the main GVC hubs Japan, USA, and Germany to exhibit strong forward linkages into GVCs.
+In particular low and middle-income countries without raw materials such as Cambodia, Mexico, or Turkey have in contrast very weak linkages and have not been able to strengthen them significantly between 1995 and 2011.
+
+%country graph e2r
+
+\section{The role of developing economies: New trends and patterns in GVCs}\label{sec:news}
+
+The central advantage of our application is that we have new indicators for a new set of countries.
+In particular, the OECD ICIO database extends the WIOD database by the following 21 regions: Argentina, Brunei Darussalam, Cambodia, Chile, Colombia, Costa Rica, Croatia, Hong Kong, Iceland, Israel, Malaysia, Norway, New Zealand, Philippines, Saudi Arabia, Singapore, Thailand, Tunisia, Vietnam, South Africa, and Switzerland.
+This means that in particular the coverage low and middle income countries has increased considerably which allows us to analyse the GVC integration in developing economies in a much more detailed fashion than before.
+In particular, we will merge a subset of the new countries into the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) and analyse the development of their GVC participation\footnote{CSA covers Argentina, Chile, Colombia, and Costa Rica; SEA covers Cambodia, Malaysia, Philippines, Thailand, and Vietnam; AFR covers South Africa and Tunisia.}.
+
+\subsection{General trends in the participation of developing economies}
+
+Regarding the integration of low- and middle-income countries, \citet{rojoguno12a} have observed that per capita income is only a weak predictor for GVC integration due to the heterogeneity of developing economies in terms of size, industrial structure and location.
+In Table \ref{tab:gvc} we see that the average integration measured by either $i2e$ or $e2r$ does not vary strongly between income groups defined by the World Bank classification at the beginning of the sample period in 1995.
+High-income economies have slightly stronger forward linkages but lower backward linkages which implies that their exports contain more domestic value added.
+Developing economies could thus try to upgrade their GVC integration by increasing domestic content.
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{GVC integration by income}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Country group & \multicolumn{2}{c}{\textit{i2e}} & \multicolumn{2}{c}{\textit{e2r}} \\
+          & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+              \midrule
+    Low/Lower middle & 23.46\% & 48.22\% & 20.35\% & 38.58\% \\
+    High & 22.64\% & 41.84\% & 21.85\% & 29.50\% \\
+    \bottomrule
+    \end{tabular}
+  \label{tab:gvc}
+\end{table}
+
+Looking at the development over time, it is striking that the rise of GVC integration is increasingly driven by developing countries.
+The growth of both $i2e$ or $e2r$ has been much more pronounced in these economies as can be seen in Table \ref{tab:gvc}.
+The $i2e$ share of countries classified as low- or lower middle-income in total $i2e$ has increased from 9\% in 1995 to 24\% in 2011.
+Similarly, the $e2r$ share has increased from 9\% to 23\%.
+
+Moreover, low- and lower middle-income countries do not only sell and source more from GVCs but they are also increasingly on the other side of the transaction.
+The $i2e$ share sourced from low- and lower middle-income countries has risen from 17\% to 33\% and the $e2r$ share re-exported has expanded from 15\% to 28\%.
+Thus, developing countries have a large stake in GVCs and have moved from the periphery into the centre of these production networks.\footnote{However, we will see that GVC integration differs significantly among developing countries.}
+
+%insert graph for 17 to 33 and 15 to 28
+
+In a next step we zoom in and analyse the GVC participation of developing economies more closely with the help of the WWZ decomposition.
+As described in section \ref{sub:wwz}, \citet{zhwaetal13} show the structure and changes in the structure of domestic and foreign content in exports inform us on a country's movement along the value chain.
+In particular, $i2e$ consists of foreign value added in final goods exports ($fva\_fin$), intermediate goods exports ($fva\_int$), and double counting ($fdc$).
+On average, Table \ref{tab:wwz} shows that low- and lower middle-income countries have a higher share of $fva\_fin$ in $i2e$ (42\%) than high-income economies (39\%).
+This is in line with a specialisation of developing economies in downstream assembly tasks.
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{WWZ decomposition results by income}
+    \begin{tabular}{lrrrrr}
+    \toprule
+    Country group & \multicolumn{1}{c}{\textit{fva\_fin}} & \multicolumn{1}{c}{\textit{fva\_inter}} & \multicolumn{1}{c}{\textit{dva\_fin}} & \multicolumn{1}{c}{\textit{dva\_inter}} & \multicolumn{1}{c}{\textit{rdv}} \\
+    \midrule
+    Low/Lower middle & 42.07\% & 57.93\% & 44.09\% & 54.73\% & 1.18\% \\
+    High  & 39.38\% & 60.62\% & 40.73\% & 56.85\% & 2.42\% \\
+    \bottomrule
+    \end{tabular}
+  \label{tab:wwz}
+\end{table}
+
+However, a shift from foreign content in final goods to intermediate goods and double counted value would be indicative of moving up the value chain.
+For low- and lower middle-income countries, we indeed find that the share of $fva\_fin$ in $i2e$ has fallen by about 4\%.
+This gain accrues to the double counting part which goes up by 6\%.
+This means that production has become more fragmented and that developing economies increasingly occupy more upstream stages.
+
+A similar exercise can be done for the domestic value added embodied in exports.
+The domestic value added of high-income countries tends to be dominated by intermediate goods (57\%) while for instance low-income countries only achieve a value of 55\%.
+We obtain the same information when we look at the share of domestic value added that eventually returns home.
+Here the value for high-income countries (2.42\%) is more than twice as high than its low- and lower-middle income counterpart (1.18\%), which indicates that high-income countries are located upstream in the value chain using developing economies for assembly.
+Here as well, the data shows though that developing economies have improved their position over time.
+The amount of domestic value added returning home has tripled from 1995 to 2011 and the share of final goods has decreased by more than 5\%.
+
+%insert graph on change over time
+
+Thus, overall we get a clear picture that while developing economies are still positioned relatively more downstream in the value chain, they have succeeded to move up over the past two decades.
+
+\subsection{Revealing new trends in the participation of developing economies}
+
+The trends described in the previous section inform us on the average performance of developing countries but they might hide considerable heterogeneity among these countries.
+Therefore, we focus in this section on selected countries in the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) described above, for which data has not been available previously.\\\
+
+\textit{South East Asia}\\
+The SEA economies for which data is newly available are Cambodia, Hong Kong, Malaysia, Philippines, Singapore, Thailand, and Vietnam.
+Since Singapore and Hong Kong are special cases due to their per capita income and size, we focus on Cambodia, Malaysia, Philippines, Thailand, and Vietnam.
+
+The two basic indicators of these countries, $i2e$ and $e2r$, presented in Table \ref{tab:seagvc} show that all five countries are primarily integrated into GVCs through backward linkages but in particular the Philippines have increased their forward linkages over the past two decades considerably.
+It also stands out that Cambodia and Vietnam have very low $e2r$ values suggesting a strong specialisation in low value added tasks located downstream in the chain.
+To obtain more detailed information on how these countries engage in GVCs we need however more disaggregated indicators.
+
+The WWZ decomposition provides us with the necessary tools.
+We can see in Table \ref{tab:seawwz} that according to their high $fva\_fin$ values Cambodia and to a lesser extent Vietnam indeed perform mostly downstream tasks with typically low value added whereas Malaysia, Thailand, and the Philippines are higher positioned in the value chain exhibiting much lower $fva\_fin$ and $dva\_fin$ but higher $rdv$ values.
+The latter set of countries thus have a similar GVC integration structure as Indonesia but still are behind Korea and Taiwan.
+
+When we look at the change over time from 1995 to 2011, we see that Cambodia has actually moved into assembly with an increase of $fva\_fin$ of 35.2\%.
+This is in stark contrast to the remaining SEA countries which all achieved to move up the value chain.
+In particular, Vietnam is on a good path with the highest decline of $fva\_fin$ and might soon catch up with its local competitors regarding its position in GVCs.
+For Cambodia, on the other hand, this means that GVCs offer a major untapped potential for future growth.
+If the country manages to introduce more GVC-friendly policies it can leverage its location close to the GVC hubs China and Japan to continue on its successful growth path.\\
+
+\begin{table}[!h]\small
+  \centering
+  \caption{GVC integration of SEA countries}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Country & \multicolumn{2}{c}{\multirow{1}[0]{*}{$i2e$}} & \multicolumn{2}{c}{\multirow{1}[0]{*}{$e2r$}} \\
+    \multicolumn{1}{c}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+    \midrule
+    Cambodia & 39.4\% & 90.7\% & 8.4\% & -11.9\% \\
+    Malaysia & 44.3\% & 37.1\% & 13.9\% & 10.2\% \\
+    Philippines & 29.6\% & -20.7\% & 22.6\% & 105.0\% \\
+    Thailand & 36.9\% & 64.3\% & 13.1\% & 20.2\% \\
+    Vietnam & 38.3\% & 66.1\% & 10.6\% & 5.3\% \\
+    \bottomrule
+    \end{tabular}
+  \label{tab:seagvc}
+\end{table}
+
+\begin{table}[!h]\small
+  \centering
+  \caption{WWZ decomposition results for SEA countries}
+  \hspace*{-2.7cm}
+    \begin{tabular}{lrrrrrrrrrr} 
+    \toprule
+    \multicolumn{1}{l}{\multirow{1}[0]{*}{Country}} & \multicolumn{2}{c}{$fva\_fin$} & \multicolumn{2}{c}{$fva\_inter$} & \multicolumn{2}{c}{$dva\_fin$} & \multicolumn{2}{c}{$dva\_inter$} & \multicolumn{2}{c}{$rdv$} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} &
+\multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+  \midrule
+    Cambodia & 68.1\% & 35.2\% & 31.9\% & -35.7\% & 64.5\% & 26.8\% & 35.5\% & -27.5\% & 0.0\% & -29.7\% \\
+    Malaysia & 39.3\% & -9.0\% & 60.7\% & 6.3\% & 40.8\% & -4.5\% & 58.9\% & 3.4\% & 0.4\% & -21.5\% \\
+    Philippines & 35.5\% & -21.7\% & 64.5\% & 16.0\% & 38.9\% & -19.1\% & 60.9\% & 16.0\% & 0.2\% & 18.2\% \\
+    Thailand & 41.4\% & -12.9\% & 58.6\% & 11.3\% & 47.4\% & -14.6\% & 52.3\% & 17.7\% & 0.3\% & 20.0\% \\
+    Vietnam & 47.1\% & -22.6\% & 52.9\% & 30.0\% & 55.0\% & -9.0\% & 44.8\% & 12.7\% & 0.1\% & 103.4\% \\
+    \bottomrule
+    \end{tabular}
+  \label{tab:seawwz}
+\end{table}
+
+\textit{Central and South America}\\
+The newly available CSA economies comprise Argentina, Chile, Colombia, and Costa Rica compared to previously only Mexico and Brazil.
+What stands out from looking at the standard GVC indicators presented in Table \ref{tab:csagvc} is that CSA is on average less integrated into GVCs than SEA.
+In particular, Argentina and Colombia have both very low backward and forward linkages highlighting the role of remoteness and sound policies as drivers of GVC integration.
+This is also mirrored in the fact that Chile and Costa Rica exhibit much higher GVC participation rates; albeit still below the SEA countries.
+These countries perform relatively well in several measures capturing a country's policy environment such as the World Bank's Doing Business Indicators or World Governance Indicators and, in the case of Costa Rica, are relatively closer to the North American GVC centre encompassing the USA, Canada, and Mexico.
+
+Focusing therefore on Costa Rica and Chile, we observe in Table \ref{tab:csawwz} that Chile's GVC integration structure resembles already the structure of high income countries.
+The major part of the country's integration is through intermediates as shown by the high $fva_inter$ and $dva_inter$ shares (78\% and 75\% respectively).
+However, the share of returned domestic value ($rdv$) is still much lower than the high-income average of 2.4\% and thus indicates that Chile is still in the process of catching up.
+
+Costa Rica on the other hand possesses the typical GVC integration structure of lower middle-income economies with high $fva_fin$ and $dva_fin$ shares and a very small $rdv$ value of 0.02\%.
+Comparing the country to SEA, its structure resembles most closely the GVC integration of Vietnam.
+This comparison holds also when we look at Costa Rica's development over time, where we see a rapid expansion of $fva_inter$, $dva_inter$, and $rdv$ shares.
+The country is thus successfully moving up the value chain.
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{GVC integration of CSA countries}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Country & \multicolumn{2}{c}{\multirow{1}[0]{*}{$i2e$}} & \multicolumn{2}{c}{\multirow{1}[0]{*}{$e2r$}} \\
+    \multicolumn{1}{c}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+    \midrule
+    Argentina & 13.4\% & 154.9\% & 13.4\% & 19.4\% \\
+    Chile & 20.0\% & 44.8\% & 26.4\% & 35.4\% \\
+    Colombia & 13.2\% & 15.2\% & 17.0\% & 45.5\% \\
+    Costa Rica & 29.0\% & 21.1\% & 16.0\% & 60.8\% \\
+    \bottomrule
+\end{tabular}
+  \label{tab:csagvc}
+\end{table}
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{WWZ decomposition results for CSA countries}
+  \hspace*{-2.7cm}
+    \begin{tabular}{lrrrrrrrrrr} 
+    \toprule
+    \multicolumn{1}{l}{\multirow{1}[0]{*}{Country}} & \multicolumn{2}{c}{$fva\_fin$} & \multicolumn{2}{c}{$fva\_inter$} & \multicolumn{2}{c}{$dva\_fin$} & \multicolumn{2}{c}{$dva\_inter$} & \multicolumn{2}{c}{$rdv$} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} &
+\multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+  \midrule
+    Argentina & 51.15\% & -6.33\% & 48.85\% & 7.95\% & 51.92\% & -5.71\% & 47.90\% & 7.01\% & 0.18\% & 49.61\% \\
+    Chile & 22.23\% & -22.29\% & 77.77\% & 8.54\% & 24.61\% & -23.44\% & 75.25\% & 10.03\% & 0.14\% & 81.25\% \\
+    Colombia & 39.41\% & -19.09\% & 60.59\% & 15.30\% & 39.32\% & -33.76\% & 60.55\% & 32.06\% & 0.12\% & 25.56\% \\
+    Costa Rica & 45.99\% & -11.17\% & 54.01\% & 11.29\% & 50.97\% & -17.88\% & 49.01\% & 24.97\% & 0.02\% & 43.05\% \\
+   \bottomrule
+    \end{tabular}
+  \label{tab:csawwz}
+\end{table}
+
+\textit{Africa}\\
+To conclude, we turn to Africa.
+GVC data on Africa is scarce and typically it is assumed that integration levels are low.
+However, the newly available OECD data comprises with Tunisia and South Africa two interesting and unique cases.
+Tunisia and South Africa offer two optimal case studies given their relatively stable political situation and relatively high degree of industrialisation.
+Unlike many other African they do thus fulfil the basic requirements for GVC integration.
+
+In line with this we find that in fact Tunisia has relatively high integration levels.
+Its integration structure is very similar in both intensity, structure, and trend to Costa Rica and Vietnam.
+This shows that especially North Africa with its proximity to the European GVC hub can benefit considerably from GVCs.
+
+South Africa is a different case since it is located far from most production networks and focuses primarily on raw materials.
+As a result, the country's integration levels are fairly low and more similar to Argentina and Colombia.
+Nevertheless, it is likely that it has benefitted from the boom in commodities caused by the rise of GVCs.
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{GVC integration of AFR countries}
+    \begin{tabular}{lrrrr}
+    \toprule
+    Country & \multicolumn{2}{c}{\multirow{1}[0]{*}{$i2e$}} & \multicolumn{2}{c}{\multirow{1}[0]{*}{$e2r$}} \\
+    \multicolumn{1}{c}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+    \midrule
+    South Africa & 21.3\% & 61.4\% & 19.9\% & 16.4\% \\
+    Tunisia & 32.1\% & 35.6\% & 13.2\% & 33.1\% \\
+\bottomrule
+\end{tabular}
+  \label{tab:afrgvc}
+\end{table}
+
+\begin{table}[htbp]\small
+  \centering
+  \caption{WWZ decomposition results for AFR countries}
+  \hspace*{-2.7cm}
+    \begin{tabular}{lrrrrrrrrrr} 
+    \toprule
+    \multicolumn{1}{l}{\multirow{1}[0]{*}{Country}} & \multicolumn{2}{c}{$fva\_fin$} & \multicolumn{2}{c}{$fva\_inter$} & \multicolumn{2}{c}{$dva\_fin$} & \multicolumn{2}{c}{$dva\_inter$} & \multicolumn{2}{c}{$rdv$} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} &
+\multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+  \midrule
+South Africa & 48.76\% & -11.76\% & 51.24\% & 13.56\% & 54.43\% & -14.60\% & 45.49\% & 21.41\% & 0.08\% & 7.07\% \\
+    Tunisia & 45.09\% & -14.97\% & 54.91\% & 15.19\% & 56.62\% & -4.47\% & 43.10\% & 5.56\% & 0.28\% & 147.59\% \\
+\bottomrule
+    \end{tabular}
+  \label{tab:afrwwz}
+\end{table}
 
 \section{Conclusion}\label{sec:conclusion}
+GVCs are a major new factor in international trade.
+International production networks span across many countries and affect many industries while changing the way trade impacts domestic economies.
+This development requires new data and new statistics that appropriately capture countries' integration into GVCs.
+In this paper, we make use both such novelties in terms of data and statistics by applying a novel gross export decomposition methodology to a new expanded dataset.
+
+More precisely, we apply the Wang-Wei-Zhu decomposition based on \citet{zhwaetal13} and \citet{rokoetal14} to a new set of Inter-Country Input-Output tables built by the OECD.
+The advantage is that twofold.
+Firstly, the WWZ decomposition allows to analyse the structure of regions' GVC integration in addition to the intensity measures provided by previous decompositions.
+Secondly, the new OECD ICIOs cover a much larger fraction of developing economies than previous ICIOs.
+This allows us to get a better understanding of how these countries integrate into GVCs.
+
+We find that many ideas based on previous anecdotal evidence can be confirmed by the data.
+In particular, there is a central difference in the structure of high-income economies' integration into GVCs when compared to developing economies when it comes to their position in GVCs.
+If we abstract from primary sectors, high-income economies are typically positioned more upstream in the value chain which can be seen from the concentration of their value added in intermediate goods exports.
+In addition, they also serve as market of final demand which can be seen from their relatively high share of exported domestic value added returning home eventually for final consumption.
+
+Developing economies, on the other hand, tend to be positioned more downstream which can be deduced from the concentration of their GVC participation in final goods exports and the fact that their forward linkages and returning domestic value added tend to be low.
+These two stylised facts suggest that high-income economies use GVCs to outsource low value added downstream production stages and before reimporting the final goods.
+However, when looking at the development over time it is positive to see that many developing economies have succeeded in moving up the value chain and that the general trends point in the right direction.
+
+Finally, we use the new data to look at selected low- and middle income economies in three different regions, namely South-East Asia, Latin America and the Caribbean, and Africa.
+South-East Asia has as expected the highest levels of GVC integration while we observe more heterogeneity in Latin America and the Caribbean where especially Chile and Costa Rica perform well.
+In Africa, we find that Tunisia has developed backward linkages into GVCs, which shows that Northern Africa has the potential to become part of the European GVC network.
+
+Overall, we show that low- and middle-income countries have become an integral part of GVCs and are increasingly the driver of their expansion.
+In addition, they increasingly succeed in moving into higher value added stages of the production networks.
+While the exact implications of integration into GVCs are still being researched, it is clear that they offer significant potential for industrialisation and growth and therefore countries like the Philippines, Costa Rica, or Tunisia are in a good position and can serve as examples for comparable countries.
 

### 2016-05-19 R/last_minute_stuff.R
add countries i2e

@@ -7,9 +7,10 @@ library(ggvis)
 library(dplyr)
 
 # load data
-sector_desc <- read_excel(path = 'data/sector_desc.xlsx')
-wwz         <- read_excel(path = 'data/graphs.xlsx')
-pdc         <- read_excel(path = 'data/graphs2.xlsx')
+country_desc <- read_excel(path = 'data/country_stats.xlsx')
+sector_desc  <- read_excel(path = 'data/sector_desc.xlsx')
+wwz          <- read_excel(path = 'data/graphs.xlsx')
+pdc          <- read_excel(path = 'data/graphs2.xlsx')
 
 # i2e plot top 6
 sector_desc %>%
@@ -68,6 +69,70 @@ sector_desc %>%
   arrange(e2r) %>%
   filter(e2r > 0.30 | e2r < 0.13) %>%
   ggvis(x=~sector, y=~e2r) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+
+# i2e plot top 6
+country_desc %>%
+  group_by(country) %>%
+  summarise(i2e=fva_sha) %>%
+  arrange(i2e) %>%
+  filter(i2e > 0.419) %>%
+  filter(country != 'irl') %>%
+  ggvis(x=~country, y=~i2e) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# i2e plot bottom 6
+country_desc %>%
+  group_by(country) %>%
+  summarise(i2e=fva_sha) %>%
+  arrange(i2e) %>%
+  filter(i2e < 0.128) %>%
+  ggvis(x=~country, y=~i2e) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+
+# i2e plot top 6 and bottom 6
+country_desc %>%
+  group_by(country) %>%
+  summarise(i2e=fva_sha) %>%
+  arrange(i2e) %>%
+  filter(i2e > 0.419 | i2e < 0.128) %>%
+  filter(country != 'irl') %>%
+  ggvis(x=~country, y=~i2e) %>%
+  layer_bars() %>%
+  scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
+
+# e2r plot top 6
+country_desc %>%
+  group_by(country) %>%
[…]

### 2016-05-19 R/last_minute_stuff.R
add countries e2r

@@ -94,7 +94,6 @@ country_desc %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
 
-
 # i2e plot top 6 and bottom 6
 country_desc %>%
   group_by(country) %>%
@@ -109,9 +108,9 @@ country_desc %>%
 # e2r plot top 6
 country_desc %>%
   group_by(country) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dva_sha) %>%
   arrange(e2r) %>%
-  filter(e2r > 0.30) %>%
+  filter(e2r > 0.265) %>%
   ggvis(x=~country, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
@@ -119,9 +118,9 @@ country_desc %>%
 # e2r plot bottom 6
 country_desc %>%
   group_by(country) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dva_sha) %>%
   arrange(e2r) %>%
-  filter(e2r < 0.13) %>%
+  filter(e2r < 0.14404) %>%
   ggvis(x=~country, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
@@ -131,7 +130,7 @@ country_desc %>%
   group_by(country) %>%
   summarise(e2r=fvax_s/exp) %>%
   arrange(e2r) %>%
-  filter(e2r > 0.30 | e2r < 0.13) %>%
+  filter(e2r > 0.265 | e2r < 0.14404) %>%
   ggvis(x=~country, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)

### 2016-05-19 R/last_minute_stuff.R
use dva for e2r

@@ -128,7 +128,7 @@ country_desc %>%
 # e2r plot top 6 and bottom 6
 country_desc %>%
   group_by(country) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dva_sha) %>%
   arrange(e2r) %>%
   filter(e2r > 0.265 | e2r < 0.14404) %>%
   ggvis(x=~country, y=~e2r) %>%

### 2016-05-19 Draft.tex
updated draft

@@
 
-In this section we look into stylised facts on GVC integration that are typically based on smaller samples with our extensive OECD ICIO dataset.
+In this section we look with our extensive OECD ICIO dataset into stylised facts on GVC integration that are typically based on smaller samples.
 We start by examining the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward and backward linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011.
@@
 
-\begin{figure}
-\centering
-\label{fig:i2e_t}
-\includegraphics[scale=0.4]{i2epercentage.png}
-\caption{The development of GVC integration over time}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\label{fig:i2e_t}
+%\includegraphics[width=7cm]{i2epercentage.png}
+%\caption{The development of GVC integration over time}
+%\end{figure}
 
@@
 
-%pdc change over time figure
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\vspace{0.8cm}
+\centering
+\includegraphics[width=\textwidth]{i2epercentage.png}
+\caption{The development of GVC integration over time}
+\label{fig:i2e_t}
+\end{minipage}\hfill
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{pdcII.png}
+\caption{The development of double counted trade over time}
+\label{fig:pdc_t}
+\end{minipage}
+\end{figure}
 
@@
 
-% sectoral graph i2e
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{industries_i2e_top6.png}
+\end{minipage}\hfill
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{industries_i2e_bottom6.png}
+\end{minipage}
+\caption{Sectoral backward linkages into GVCs - Top and bottom 6.}
+\label{fig:i2e_s}
+\end{figure}
 
@@
 In that respect, it is also interesting to note that despite the low absolute $i2e$ shares, it is in services where a lot of the growth in $i2e$ has taken place.
-As can be seen in Figure \ref{fig:i2e_st}, five out of the six sectors with the highest growth in $i2e$ shares are services sectors.
+Five out of the six sectors with the highest growth in $i2e$ shares are services sectors.
 
-% sectoral graph e2r
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{industries_e2r_top6.png}
+\end{minipage}\hfill
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{industries_e2r_bottom6.png}
+\end{minipage}
+\caption{Sectoral forward linkages into GVCs - Top and bottom 6.}
+\label{fig:e2r_s}
+\end{figure}
 
@@
 
-%country graph i2e
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{country_i2e_top_6.png}
+\end{minipage}\hfill
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{countries_i2e_bottom_6.png}
+\end{minipage}
+\caption{Countries' backward linkages into GVCs - Top and bottom 6.}
+\label{fig:i2e_k}
+\end{figure}
 
@@
 
-%country graph e2r
+\begin{figure}
+\centering
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{country_e2r_top_6.png}
+\end{minipage}\hfill
+\begin{minipage}{0.45\textwidth}
+\centering
+\includegraphics[width=\textwidth]{countries_e2r_bottom_6.png}
+\end{minipage}
+\caption{Countries' forward linkages into GVCs - Top and bottom 6.}
+\label{fig:e2r_k}
+\end{figure}
 

### 2016-05-19 R/last_minute_stuff.R
fix e2r industries

@@ -45,9 +45,9 @@ sector_desc %>%
 # e2r plot top 6
 sector_desc %>%
   group_by(sector) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dvar_s/exp) %>%
   arrange(e2r) %>%
-  filter(e2r > 0.30) %>%
+  filter(e2r > 0.3) %>%
   ggvis(x=~sector, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
@@ -55,9 +55,9 @@ sector_desc %>%
 # e2r plot bottom 6
 sector_desc %>%
   group_by(sector) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dvar_s/exp) %>%
   arrange(e2r) %>%
-  filter(e2r < 0.13) %>%
+  filter(e2r < 0.108) %>%
   ggvis(x=~sector, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)
@@ -65,9 +65,9 @@ sector_desc %>%
 # e2r plot top 6 and bottom 6
 sector_desc %>%
   group_by(sector) %>%
-  summarise(e2r=fvax_s/exp) %>%
+  summarise(e2r=dvar_s/exp) %>%
   arrange(e2r) %>%
-  filter(e2r > 0.30 | e2r < 0.13) %>%
+  filter(e2r > 0.30 | e2r < 0.108) %>%
   ggvis(x=~sector, y=~e2r) %>%
   layer_bars() %>%
   scale_numeric("y", domain = c(0, 0.5), nice = FALSE)

### 2016-05-19 R/last_minute_stuff.R
fix title fva dva

@@ -162,13 +162,15 @@ wwz2 %>%
   layer_lines() %>%
   layer_lines(x=~year, y=~fva_inter, stroke='fva_inter') %>%
   layer_lines(x=~year, y=~dva_fin,   stroke='dva_fin')   %>%
-  layer_lines(x=~year, y=~dva_inter, stroke='dva_inter')
+  layer_lines(x=~year, y=~dva_inter, stroke='dva_inter') %>%
+  add_axis('y', title = 'pdc_t')
+
 
 pdc %>%
   ggvis(x=~year) %>%
   layer_lines(y=~pdc_t) %>%
   add_axis('y', orient = 'left', title = 'pdc_t') %>%
-  add_axis('y', 'ydiv' , orient = 'right',
+  add_axis('y', 'ydiv', orient = 'right',
            title= 'pdc_exp', grid=F, title_offset = 50,
            properties = axis_props(labels = list(fill = 'red')) ) %>%
   layer_lines( prop('y', ~pdc_exp, scale='ydiv'), stroke:='red' )

### 2016-05-19 R/last_minute_stuff.R
fix title e2r i2e

@@ -140,7 +140,6 @@ names(wwz1) <- c('year', 'e2rL-LM', 'i2eL-LM')
 wwz1$year <- as.numeric(wwz1$year)
 wwz1$`e2rL-LM` <- as.numeric(wwz1$`e2rL-LM`)
 wwz1$`i2eL-LM` <- as.numeric(wwz1$`i2eL-LM`)
-wwz1$`e2rL-LM` <- round(wwz1$`e2rL-LM`)
 
 wwz2 <- wwz[11:17,]
 names(wwz2) <- wwz[10,]
@@ -155,7 +154,8 @@ wwz2$rdv       <- as.numeric(wwz2$rdv)
 wwz1 %>%
   ggvis(x=~year, y=~`e2rL-LM`, stroke='e2r') %>%
   layer_lines() %>%
-  layer_lines(x=~year, y=~`i2eL-LM`, stroke='i2e')
+  layer_lines(x=~year, y=~`i2eL-LM`, stroke='i2e') %>%
+  add_axis('y', title = 'e2r / i2e')
 
 wwz2 %>%
   ggvis(x=~year, y=~fva_fin, stroke='fva_fin') %>%
@@ -163,7 +163,8 @@ wwz2 %>%
   layer_lines(x=~year, y=~fva_inter, stroke='fva_inter') %>%
   layer_lines(x=~year, y=~dva_fin,   stroke='dva_fin')   %>%
   layer_lines(x=~year, y=~dva_inter, stroke='dva_inter') %>%
-  add_axis('y', title = 'pdc_t')
+  layer_lines(x=~year, y=~rdv,       stroke='rdv')       %>%
+  add_axis('y', title = 'WWZ decomposition indicators')
 
 
 pdc %>%

### 2016-05-19 R/last_minute_stuff.R
add genius plot

@@ -163,8 +163,12 @@ wwz2 %>%
   layer_lines(x=~year, y=~fva_inter, stroke='fva_inter') %>%
   layer_lines(x=~year, y=~dva_fin,   stroke='dva_fin')   %>%
   layer_lines(x=~year, y=~dva_inter, stroke='dva_inter') %>%
-  layer_lines(x=~year, y=~rdv,       stroke='rdv')       %>%
-  add_axis('y', title = 'WWZ decomposition indicators')
+  add_axis('y', orient='left', title = 'WWZ decomposition indicators')  %>%
+  add_axis('y', 'ydiv', orient = 'right',
+           title= 'rdv', grid=F, title_offset = 50,
+           properties = axis_props(labels = list(fill = 'purple')) ) %>%
+  layer_lines( prop('y', ~rdv, scale='ydiv'), stroke='rdv' )
+
 
 
 pdc %>%

### 2016-05-19 Draft.tex
final version

@@
 \usepackage{graphicx}
+\usepackage{caption}
 
@@
 A central step towards a more in-depth analysis of GVCs has been laid by \citet{rokoetal14} and \citet{zhwaetal13} who show that it is necessary to go beyond deriving origins of value added to examine production sharing comprehensively.
-They split exports into different categories and calculate metrics of how often goods cross borders.
+They split goods into different categories and calculate metrics of how often these goods cross borders.
 This enables them to derive measures of GVC length but also allows to investigate how individual countries are integrated into GVCs.
@@
 
-However, these contributions typically have two shortcomings.
+However, these contributions typically have one of two shortcomings.
 Firstly, most evidence is based on data from high-income countries.
@@
 In addition, we find that high-income countries typically are the starting and end points of GVCs in that they provide upstream inputs and then serve eventually again as demand markets for the final products.
-Low- and middle-income countries, on the other hand, are rather specialised in downstream activities such as assembly and export typically less domestic value added.
-However, we also observe that developing economies move more and more upstream along the value chain and should thus be able to generate more gains from their GVC participation.
+Low- and middle-income countries, on the other hand, are more specialised in downstream activities such as assembly and export typically less domestic value added.
+However, we observe that developing economies have begun to move upstream along the value chain and out of pure assembly occupying a wider set stages.
+This should allow them to generate more gains from GVC participation.
 
@@
 
-The idea is that the production of industry \textit{i} creates value added in industry $i$ itself, a direct contribution, but also in industries that supply $i$ with inputs, an indirect contribution.
+The idea is that the production of industry \textit{i} of country \textit{k} creates value added in industry $i$ itself, a direct contribution, but also in industries \textit{j} from \textit{k} or other countries \textit{l} that supply $i$ with inputs, an indirect contribution.
 Since these industries themselves rely on inputs, $i$'s production sets several rounds of indirect value added creation in motion that can mathematically be expressed as:
@@
 \textit{VB} gives thus so called value added multipliers, which denote the amount of value added that the production of an industry's 1\$ of output or exports brings about in all other industries.
-If we post-multiply $VB$ with exports, we get a matrix with the value added origins of each industry's exports.
+If we post-multiply $VB$ with exports, we get a matrix, $VAE$, with the elements being the value added origins of each industry's exports, $vae_{ikjl}$.
 
-This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures, namely foreign value added in exports (Hummels et al.'s Vertical Specialisation) and domestic value added in foreign exports (Hummels et al.'s Vertical Specialisation 1).
+This basic decomposition has been widely used in GVC analysis since it allows the calculation of two informative GVC participation measures.
+Firstly, a backward linkage indicator that is given by the import content of exports, $i2e$, (\citet{dahuetal01}`s Vertical Specialisation) and calculated as follows:
+\begin{equation}
+i2e_{ik} = \frac{\sum_{l}{\sum_{j}vae_{jlik}}}{exports_{ik}} ,
+\end{equation}
+Secondly, a forward linkage indicator - \emph{e2r} (domestic content in foreign (re-)exports) - which is given by:
+\begin{equation}\label{eq:dvar}
+{e2r}_{ik} = \frac{\sum_{l}{\sum_{j} vae_{ikjl}}}{exports_{ik}} ,
+\end{equation}
+where \(l\  \neq k\).
+
 These indicators can tell us how much a country is integrated into GVCs and if it acts mainly as a supplier or a user of foreign value added.
@@
 absorbed abroad).
-This splits the exports, $E$, of industry $l$ in country $k$ into sixteen different parts broadly differentiated into the four boroad categories domestic value added absorbed abroad, domestic value added returning home, foreign value added, and purely double counted terms:
+This splits the exports, $E$, of industry $l$ in country $k$ into sixteen different parts broadly differentiated into the four broad categories domestic value added absorbed abroad, domestic value added returning home, foreign value added, and purely double counted terms:
 \begin{align}\label{eq:wwz}
@@
 \item Domestic value added in final exports (\textit{DVA\_FIN}, T1)
-\item Domestic value added in intermediate exports
+\item Domestic value added in intermediate exports (\textit{DVA\_INT}, T2-5)
 \begin{itemize}
@@
 \end{itemize}
-\item Domestic value added returning home (\textit{RDV\_B}, T6-8)
+\item Domestic value added returning home (\textit{RDV}, T6-8)
 \begin{itemize}
@@
 
+For the analysis, we use $dva\_fin$, $fva\_fin$, $rdv$, $pdc$ and the two aggregate measures $dva_inter$ combining $dva_int$ and $ddc$ as well as $fva_inter$ combining $fva_int$ and $fdc$.
+This collapses the indicator to a intuitive and manageable amount.
+
 The advantage of such a detailed decomposition is that these new indicators can inform us on how countries integrate into GVCs while the basic Leontief decomposition mainly informs us on the intensity of integration.
 High amounts of foreign value added in final goods exports are, for instance, suggestive of a specialisation in downstream tasks that add little value to a good, such as assembly.
-High amounts of domestic value added in intermediate exports, on the other hand, are evidence of a more upstream specialisation in tasks that aaa a lot of value, such as business services.
+High amounts of domestic value added in intermediate exports, on the other hand, are evidence of a more upstream specialisation in tasks that add a lot of value, such as business services.
 By tracking these two variables over time we can then see which countries have succeeded in moving up the value chain.
+We will explain the indicators in more detail in combination with the decomposition results to facilitate the understanding.
 
-The high resolution of the WWZ decomposition does, however, not mean that the Leontief decomposition does not contain valuable information at all.
+Finally, it is necessary to point out that the high resolution of the WWZ decomposition does not mean that the Leontief decomposition does not contain valuable information at all.
 In fact, we exploit the decomposition of exports into source industry and source country by calculating variants of the standard indicators based on different characteristics.
@@
 
-In this section we look with our extensive OECD ICIO dataset into stylised facts on GVC integration that are typically based on smaller samples.
-We start by examining the development of the most basic measure of GVC integration introduced by \citet{dahuetal01}, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward and backward linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011.
+In this section we reassess with our extensive OECD ICIO dataset stylised facts on GVC integration that are typically based on smaller samples.
+We start by examining the development of our most basic measure of GVC integration, namely the amount of foreign value added in exports labeled by \citet{ribajalo13} as $i2e$.\footnote{Note that at the aggregate level forward ($e2r$) and backward ($i2e$) linkages are identical and thus we only look at one of the two measures.} It captures backward linkages into value chains and shows the well-known increase in GVC integration from 1995 to 2011.
 As illustrated in Figure \ref{fig:i2e_t}, the value of $i2e$ has grown by approximately 350\% and by 35\% as a share of total exports from around 17\% to over 23\%.
@@
 
-%\begin{figure}[h]
-%\centering
-%\label{fig:i2e_t}
-%\includegraphics[width=7cm]{i2epercentage.png}
-%\caption{The development of GVC integration over time}
-%\end{figure}
-
-Another way to examine the expansion of GVCs from 1995 to 2011 is to look at their intensity or length instead of their trade volume.
-\citet{zhwaetal13} propose to use the amount of double counted trade, $pdc$, as a proxy for GVC length since its value goes up the more back-and-forth trade occurs, which is equivalent to an increase in the number of production stages.
-They show that its value has increased for the 40 WIOD countries.
-In Figure \ref{fig:pdc_t}, we observe similarly that $pdc$ as a share of total exports has increased over the examined period by 73\% and thus more than $i2e$.
+Another way to examine the expansion of GVCs from 1995 to 2011 is to look at their length instead of their trade volume.
+WWZ propose to use the amount of double counted trade, $pdc$, as a proxy for GVC length since its value goes up the more back-and-forth trade occurs, which is equivalent to an increase in the number of production stages.
+They show that its value has increased for 40 selected countries.
+In Figure \ref{fig:pdc_t}, we observe in our larger sample similarly that $pdc$ as a share of total exports has increased over the examined period by 73\% and thus more than $i2e$.
 Therefore, GVCs do not only channel more trade but also have become longer over time.
@@
 
-Turning from the development over time to sectoral differences in GVC integration, Figure \ref{fig:i2e_s} reveals in line with \citet{rojoguno12a} that the sectors exhibiting the highest dependency on foreign inputs in terms of $i2e$ shares are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ).
+Turning from the development over time to sectoral differences in GVC integration, Figure \ref{fig:i2e_s} reveals in line with \citet{rojoguno12a} that the sectors exhibiting the highest degree of international fragmentation in terms of $i2e$ shares are heavy manufactures such as motor vehicles (MTR), other transport equipment (TRQ) and the metal industry (MET) as well as computers and electronics (CEQ and ELQ).
 In particular, the transport equipment and electronics industry are strongly engaged in GVCs having highly international production networks.
 For instance, Apple's iPhone contains inputs from 9 to 10 countries while the Boeing 787 production spans more than 5 countries.
-The sectors can be characterised as being close to final demand or being dependent on imported raw materials.
+The sectors can be characterised as being close to final demand and producing complex differentiated goods.
 These characteristics can thus explain differences in GVC integration.
@@
 The bottom 6 industries in terms of $i2e$ shares are primary and services sectors such as agriculture (AGR), mining (MIN), R\&D and business services (BZS), or wholesale and retail trade (WRT).
-These sectors are typically located upstream in the supply chain far from final demand and have high high value added to output ratios.
+These sectors are typically located upstream in the supply chain far from final demand and have high value added to output ratios.
 
@@
 \end{minipage}
-\caption{Sectoral backward linkages into GVCs - Top and bottom 6.}
+\caption{Sectoral $i2e$ shares - Top and bottom 6.}
 \label{fig:i2e_s}
@@
 It captures the amount of domestic value added in foreign exports and thus quantifies how important domestic industries are for foreign export production.
-Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $i2e$ ranking such as mining or financial and telecommunication services (FIN and PTL).
+Here, Figure \ref{fig:e2r_s} depicts that this indicator is dominated by the same upstream industries that are at the bottom of the $i2e$ ranking such as mining or business and telecommunication services (PTL).
 This shows that these industries are also strongly engaged in GVCs but their participation is of a different type.
@@
 With respect to services, it is also indicative of the servicification of manufacturing as described by \citet{ribaetal15}.
-This means that an increasing share of manufacturing gross exports is actually value added generated in services sectors and then embedded in the intermediate exports of manufacturers.
+This means that an increasing share of manufacturing gross exports is actually value added generated in services sectors and then embedded in the intermediate goods exports of manufacturers.
 This importance of services sectors for exports cannot be seen from standard gross trade statistics and thus constitutes a major advantage of trade in value added measures.
+
 It is also indicative of a growing internationalisation of services.
-Increasingly, services are being offshored and sourced from abroad.
+More and more, services are being offshored and sourced from abroad.
 In that respect, it is also interesting to note that despite the low absolute $i2e$ shares, it is in services where a lot of the growth in $i2e$ has taken place.
@@
 \end{minipage}
-\caption{Sectoral forward linkages into GVCs - Top and bottom 6.}
+\caption{Sectoral $e2r$ shares - Top and bottom 6.}
 \label{fig:e2r_s}
@@
 Finally, when we turn to differences in GVC integration by country we can confirm the findings by \citet{ribajalo13}.
-Figure \ref{i2e_k} shows that small countries close to the major GVC hubs in Asia, Europe, and North America such as Slovakia or Malaysia have the highest average $i2e$ shares.
+Figure \ref{fig:i2e_k} shows that small countries close to the major GVC hubs in Asia, Europe, and North America have the highest average $i2e$ shares.
+Examples include Malaysia and Slovakia.
 Countries specialised in the primary sector or assembly on the other hand have very low values.
@@
 However, the development over time shows that some of the countries with the relatively low GVC integration have begun to catch up.
-For instance, Argentina, Indian and Turkey are in the top 6 when it comes to the growth of $i2e$ shares from 1995 to 2011.
+For instance, Argentina, India and Turkey are in the top 6 when it comes to the growth of $i2e$ shares from 1995 to 2011.
 
-\begin{figure}
+Driven by the sectoral statistics, we then find again that for $e2r$ the picture is reversed with raw material exporters on top.
+If we abstract from these countries we find technologically advanced countries such as Switzerland and the main GVC hubs Japan, USA, and Germany to exhibit strong forward linkages into GVCs.
+In particular low and middle-income countries without raw materials such as Cambodia, Mexico, or Turkey have in contrast very weak linkages and have not been able to strengthen them significantly between 1995 and 2011.\footnote{The full set of results for $i2e$ and $e2r$ by country, and sector can be found in the Appendix.
+Since the results of WWZ decomposition are much more detailed the results not presented here are only available from the authors upon request.}
+
+\begin{figure}[h]
 \centering
@@
 \end{minipage}
-\caption{Countries' backward linkages into GVCs - Top and bottom 6.}
+\caption{Countries' $i2e$ shares - Top and bottom 6.}
 \label{fig:i2e_k}
@@
 
-Driven by the sectoral statistics, we then find again that for $e2r$ the picture is reversed with raw material exporters on top.
-If we abstract from these countries we find technologically advanced countries such as Switzerland and the main GVC hubs Japan, USA, and Germany to exhibit strong forward linkages into GVCs.
-In particular low and middle-income countries without raw materials such as Cambodia, Mexico, or Turkey have in contrast very weak linkages and have not been able to strengthen them significantly between 1995 and 2011.
-
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \centering
-\includegraphics[width=\textwidth]{country_e2r_top_6.png}
+\includegraphics[width=\textwidth]{countries_e2r_top_6.png}
 \end{minipage}\hfill
@@
 \end{minipage}
-\caption{Countries' forward linkages into GVCs - Top and bottom 6.}
+\caption{Countries' $e2r$ shares - Top and bottom 6.}
 \label{fig:e2r_k}
@@
 The central advantage of our application is that we have new indicators for a new set of countries.
-In particular, the OECD ICIO database extends the WIOD database by the following 21 regions: Argentina, Brunei Darussalam, Cambodia, Chile, Colombia, Costa Rica, Croatia, Hong Kong, Iceland, Israel, Malaysia, Norway, New Zealand, Philippines, Saudi Arabia, Singapore, Thailand, Tunisia, Vietnam, South Africa, and Switzerland.
-This means that in particular the coverage low and middle income countries has increased considerably which allows us to analyse the GVC integration in developing economies in a much more detailed fashion than before.
-In particular, we will merge a subset of the new countries into the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) and analyse the development of their GVC participation\footnote{CSA covers Argentina, Chile, Colombia, and Costa Rica; SEA covers Cambodia, Malaysia, Philippines, Thailand, and Vietnam; AFR covers South Africa and Tunisia.}.
+This means we can not only confirm previous findings with a more representative sample but also provide several new insights.
+In particular, the OECD ICIO database extends the available list of countries in ICIOs by the following 21 regions: Argentina, Brunei Darussalam, Cambodia, Chile, Colombia, Costa Rica, Croatia, Hong Kong, Iceland, Israel, Malaysia, Norway, New Zealand, Philippines, Saudi Arabia, Singapore, Thailand, Tunisia, Vietnam, South Africa, and Switzerland.
+This means that in particular the coverage of low and middle income countries has increased considerably which allows us to analyse the GVC integration of developing economies in a much more detailed fashion than before.
 
-\subsection{General trends in the participation of developing economies}
+\subsection{General trends in the GVC participation of developing economies}
 
-Regarding the integration of low- and middle-income countries, \citet{rojoguno12a} have observed that per capita income is only a weak predictor for GVC integration due to the heterogeneity of developing economies in terms of size, industrial structure and location.
-In Table \ref{tab:gvc} we see that the average integration measured by either $i2e$ or $e2r$ does not vary strongly between income groups defined by the World Bank classification at the beginning of the sample period in 1995.
+Regarding the integration of low- and middle-income countries, \citet{rojoguno12a} have observed that per capita income is only a weak predictor for GVC integration due to the heterogeneity of economies in terms of size, industrial structure and location.
+In Table \ref{tab:gvc} we see that the average integration measured by either $i2e$ or $e2r$ does not vary strongly between income groups defined by the World Bank classification at the beginning of the sample period in 1995.\footnote{Note that in this section indicators are based only on manufacturing and services sectors to avoid spurious results stemming from primary sectors that are for technological reasons less integrated into GVCs.}
 High-income economies have slightly stronger forward linkages but lower backward linkages which implies that their exports contain more domestic value added.
-Developing economies could thus try to upgrade their GVC integration by increasing domestic content.
+Developing economies could thus try to upgrade their GVC integration by increasing domestic content in exports.
 
@@
     Country group & \multicolumn{2}{c}{\textit{i2e}} & \multicolumn{2}{c}{\textit{e2r}} \\
-          & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+          & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} \\
               \midrule
@@
   \label{tab:gvc}
+    \caption*{Data is averaged across countries, sectors and years. $\Delta$ 95-11 refers to the growth of the \textit{i2e} and \textit{e2r} values from 1995 to 2011.}
 \end{table}
@@
 The growth of both $i2e$ or $e2r$ has been much more pronounced in these economies as can be seen in Table \ref{tab:gvc}.
-The $i2e$ share of countries classified as low- or lower middle-income in total $i2e$ has increased from 9\% in 1995 to 24\% in 2011.
+In relative terms this means that the $i2e$ share of countries classified as low- or lower middle-income in total $i2e$ has increased from 9\% in 1995 to 24\% in 2011.
 Similarly, the $e2r$ share has increased from 9\% to 23\%.
@@
 Moreover, low- and lower middle-income countries do not only sell and source more from GVCs but they are also increasingly on the other side of the transaction.
-The $i2e$ share sourced from low- and lower middle-income countries has risen from 17\% to 33\% and the $e2r$ share re-exported has expanded from 15\% to 28\%.
+Figure \ref{fig:llmpartner} shows that the share of $i2e$ sourced from low- and lower middle-income countries has risen from 17\% to 33\% and the share of $e2r$ re-exported from them has expanded from 15\% to 28\%.
 Thus, developing countries have a large stake in GVCs and have moved from the periphery into the centre of these production networks.\footnote{However, we will see that GVC integration differs significantly among developing countries.}
 
-%insert graph for 17 to 33 and 15 to 28
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\textwidth]{e2r-i2e.png}
+\caption{Share of value added sourced from (i2e) or sold to (e2r) low- and lower-middle income economies for export production.}
+\label{fig:llmpartner}
+\end{figure}
 
 In a next step we zoom in and analyse the GVC participation of developing economies more closely with the help of the WWZ decomposition.
-As described in section \ref{sub:wwz}, \citet{zhwaetal13} show the structure and changes in the structure of domestic and foreign content in exports inform us on a country's movement along the value chain.
+As described in section \ref{sub:wwz}, WWZ show how the structure and changes in the structure of domestic and foreign content in exports inform us on a country's movement along the value chain.
 In particular, $i2e$ consists of foreign value added in final goods exports ($fva\_fin$), intermediate goods exports ($fva\_int$), and double counting ($fdc$).
-On average, Table \ref{tab:wwz} shows that low- and lower middle-income countries have a higher share of $fva\_fin$ in $i2e$ (42\%) than high-income economies (39\%).
+Table \ref{tab:wwz} shows that on average low- and lower middle-income countries have a higher share of $fva\_fin$ in $i2e$ (42\%) than high-income economies (39\%).
 This is in line with a specialisation of developing economies in downstream assembly tasks.
@@
   \label{tab:wwz}
+  \caption*{Data is averaged across countries, sectors and years. $fva$ variables are expressed as \% of $i2e$, $dva$ and $rdv$ variables as \% of domestic value added in total exports.}
 \end{table}
 
-However, a shift from foreign content in final goods to intermediate goods and double counted value would be indicative of moving up the value chain.
-For low- and lower middle-income countries, we indeed find that the share of $fva\_fin$ in $i2e$ has fallen by about 4\%.
+However, a shift from foreign content in final goods to intermediate goods and double counted trade value would be indicative of moving up the value chain.
+For low- and lower middle-income countries, we indeed find as shown by Figure \ref{fig:llmwwz} that the share of $fva\_fin$ in $i2e$ has fallen by about 4\%.
 This gain accrues to the double counting part which goes up by 6\%.
@@
 A similar exercise can be done for the domestic value added embodied in exports.
-The domestic value added of high-income countries tends to be dominated by intermediate goods (57\%) while for instance low-income countries only achieve a value of 55\%.
+The exported domestic value added of high-income countries tends to be dominated by intermediate goods (57\%) while low- and lower middle-income countries only achieve a value of 55\%.
 We obtain the same information when we look at the share of domestic value added that eventually returns home.
-Here the value for high-income countries (2.42\%) is more than twice as high than its low- and lower-middle income counterpart (1.18\%), which indicates that high-income countries are located upstream in the value chain using developing economies for assembly.
-Here as well, the data shows though that developing economies have improved their position over time.
+Here, the value for high-income countries (2.42\%) is more than twice as high than its low- and lower-middle income counterpart (1.18\%), which indicates that high-income countries are located upstream in the value chain using developing economies for assembly.
+However, the data shows as well that developing economies have improved their position over time.
 The amount of domestic value added returning home has tripled from 1995 to 2011 and the share of final goods has decreased by more than 5\%.
 
-%insert graph on change over time
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{fva-dva-fin-inter.png}
+\caption{Development of developing economies' WWZ decomposition indicators over time.}
+\label{fig:llmwwz}
+\end{figure}
 
@@
 The trends described in the previous section inform us on the average performance of developing countries but they might hide considerable heterogeneity among these countries. 
-Therefore, we focus in this section on selected countries in the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) described above, for which data has not been available previously.\\\
+Therefore, we will merge a subset of the newly available countries into the three regions Central and South America (CSA), South East Asia (SEA), and Africa (AFR) and analyse the development of their GVC participation country by country.
+CSA covers Argentina, Chile, Colombia, and Costa Rica; SEA covers Cambodia, Malaysia, Philippines, Thailand, and Vietnam; while AFR covers South Africa and Tunisia.\\\
 
@@
 The WWZ decomposition provides us with the necessary tools.
-We can see in Table \ref{tab:seawwz} that according to their high $fva\_fin$ values Cambodia and to a lesser extent Vietnam indeed perform mostly downstream tasks with typically low value added whereas Malaysia, Thailand, and the Philippines are higher positioned in the value chain exhibiting much lower $fva\_fin$ and $dva\_fin$ but higher $rdv$ values.
-The latter set of countries thus have a similar GVC integration structure as Indonesia but still are behind Korea and Taiwan.
+We can see in Table \ref{tab:seawwz} that according to their high $fva\_fin$ values Cambodia and to a lesser extent Vietnam indeed perform mostly downstream tasks with typically low value added whereas Malaysia, Thailand, and the Philippines are positioned higher in the value chain exhibiting much lower $fva\_fin$ and $dva\_fin$ but higher $rdv$ values.
+Comparing these results to the analysis by WWZ, we find that the latter set of countries have a similar GVC integration structure to Indonesia but still are behind more advanced nations such as Korea and Taiwan.
 
@@
     Country & \multicolumn{2}{c}{\multirow{1}[0]{*}{$i2e$}} & \multicolumn{2}{c}{\multirow{1}[0]{*}{$e2r$}} \\
-    \multicolumn{1}{c}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+    \multicolumn{1}{c}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} \\
     \midrule
@@
   \label{tab:seagvc}
+  \caption*{Data is averaged across sectors and years. $\Delta$ 95-11 refers to growth from 1995 to 2011.}
 \end{table}
@@
     \multicolumn{1}{l}{\multirow{1}[0]{*}{Country}} & \multicolumn{2}{c}{$fva\_fin$} & \multicolumn{2}{c}{$fva\_inter$} & \multicolumn{2}{c}{$dva\_fin$} & \multicolumn{2}{c}{$dva\_inter$} & \multicolumn{2}{c}{$rdv$} \\
-    \multicolumn{1}{l}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} &
-\multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{95-11} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} &
+\multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} & \multicolumn{1}{c}{Average} & \multicolumn{1}{c}{$\Delta$ 95-11} \\
   \midrule
@@
   \label{tab:seawwz}
+   \caption*{Data is averaged across sectors and years.
+$fva$ variables are expressed as \% of $i2e$, $dva$ and $rdv$ variables as \% of domestic value added in total exports.
+$\Delta$ 95-11 refers to growth from 1995 to 2011.}
 \end{table}
@@
 The newly available CSA economies comprise Argentina, Chile, Colombia, and Costa Rica compared to previously only Mexico and Brazil.
-What stands out from looking at the standard GVC indicators presented in Table \ref{tab:csagvc} is that CSA is on average less integrated into GVCs than SEA.
+What stands out from looking at the standard GVC indicators presented in Table \ref{tab:csagvc} is that CSA is on average less integrated into GVCs than SEA and other developing regions.
 In particular, Argentina and Colombia have both very low backward and forward linkages highlighting the role of remoteness and sound policies as drivers of GVC integration.
@@
 
-Focusing therefore on Costa Rica and Chile, we observe in Table \ref{tab:csawwz} that Chile's GVC integration structure resembles already the structure of high income countries.
-The major part of the country's integration is through intermediates as shown by the high $fva_inter$ and $dva_inter$ shares (78\% and 75\% respectively).
+Focusing therefore on Costa Rica and Chile, we observe in Table \ref{tab:csawwz} that Chile's GVC integration structure starts to resembles the structure of high income countries.
+The major part of the country's integration is through intermediates as shown by the high $fva\_inter$ and $dva\_inter$ shares (78\% and 75\% respectively).
 However, the share of returned domestic value ($rdv$) is still much lower than the high-income average of 2.4\% and thus indicates that Chile is still in the process of catching up.
 
-Costa Rica on the other hand possesses the typical GVC integration structure of lower middle-income economies with high $fva_fin$ and $dva_fin$ shares and a very small $rdv$ value of 0.02\%.
+Costa Rica on the other hand possesses the typical GVC integration structure of lower middle-income economies with high $fva\_fin$ and $dva\_fin$ shares and a very small $rdv$ value of 0.02\%.
 Comparing the country to SEA, its structure resembles most closely the GVC integration of Vietnam.
-This comparison holds also when we look at Costa Rica's development over time, where we see a rapid expansion of $fva_inter$, $dva_inter$, and $rdv$ shares.
+This comparison holds also when we look at Costa Rica's development over time, where we see a rapid expansion of $fva\_inter$, $dva\_inter$, and $rdv$ shares.
 The country is thus successfully moving up the value chain.
@@
   \label{tab:csagvc}
+     \caption*{Data is averaged across sectors and years. $\Delta$ 95-11 refers to growth from 1995 to 2011.}
+
 \end{table}
@@
   \label{tab:csawwz}
+     \caption*{Data is averaged across sectors and years.
+$fva$ variables are expressed as \% of $i2e$, $dva$ and $rdv$ variables as \% of domestic value added in total exports.
+$\Delta$ 95-11 refers to growth from 1995 to 2011.}
 \end{table}
@@
 However, the newly available OECD data comprises with Tunisia and South Africa two interesting and unique cases.
-Tunisia and South Africa offer two optimal case studies given their relatively stable political situation and relatively high degree of industrialisation.
+Tunisia and South Africa offer relatively stable political environments and a relatively high degree of industrialisation which makes them two optimal case studies.
 Unlike many other African they do thus fulfil the basic requirements for GVC integration.
 
-In line with this we find that in fact Tunisia has relatively high integration levels.
-Its integration structure is very similar in both intensity, structure, and trend to Costa Rica and Vietnam.
-This shows that especially North Africa with its proximity to the European GVC hub can benefit considerably from GVCs.
+In line with this, Tables \ref{tab:afrgvc} and \ref{tab:afrwwz} show that in fact Tunisia has relatively high integration levels.
+Its integration pattern is very similar in both intensity, structure, and trend to Costa Rica and Vietnam.
+This means that Tunisia is mainly integrated through backward linkages and assembly tasks but is moving up the value chain.
+This is evidence that especially North Africa with its proximity to the European GVC hub can link into and benefit from GVCs.
 
@@
 As a result, the country's integration levels are fairly low and more similar to Argentina and Colombia.
-Nevertheless, it is likely that it has benefitted from the boom in commodities caused by the rise of GVCs.
+Nevertheless, it is likely that it has benefitted from the boom in commodities caused by the rise of GVCs and the subsequent boost in global demand.
 
@@
   \label{tab:afrgvc}
+     \caption*{Data is averaged across sectors and years. $\Delta$ 95-11 refers to growth from 1995 to 2011.}
 \end{table}
@@
   \label{tab:afrwwz}
+     \caption*{Data is averaged across sectors and years.
+$fva$ variables are expressed as \% of $i2e$, $dva$ and $rdv$ variables as \% of domestic value added in total exports.
+$\Delta$ 95-11 refers to growth from 1995 to 2011.}
 \end{table}
@@
 More precisely, we apply the Wang-Wei-Zhu decomposition based on \citet{zhwaetal13} and \citet{rokoetal14} to a new set of Inter-Country Input-Output tables built by the OECD.
-The advantage is that twofold.
-Firstly, the WWZ decomposition allows to analyse the structure of regions' GVC integration in addition to the intensity measures provided by previous decompositions.
+The advantage is twofold.
+Firstly, the WWZ decomposition allows to analyse the structure of regions' GVC integration in addition to the intensity measures provided by previous decompositions leading to deeper insights into GVC integration patterns.
 Secondly, the new OECD ICIOs cover a much larger fraction of developing economies than previous ICIOs.
@@
 We find that many ideas based on previous anecdotal evidence can be confirmed by the data.
-In particular, there is a central difference in the structure of high-income economies' integration into GVCs when compared to developing economies when it comes to their position in GVCs.
+In particular, there is a central difference in the structure of high-income economies' integration into GVCs compared to developing economies when it comes to the position in GVCs.
 If we abstract from primary sectors, high-income economies are typically positioned more upstream in the value chain which can be seen from the concentration of their value added in intermediate goods exports.
@@
 Developing economies, on the other hand, tend to be positioned more downstream which can be deduced from the concentration of their GVC participation in final goods exports and the fact that their forward linkages and returning domestic value added tend to be low.
-These two stylised facts suggest that high-income economies use GVCs to outsource low value added downstream production stages and before reimporting the final goods.
-However, when looking at the development over time it is positive to see that many developing economies have succeeded in moving up the value chain and that the general trends point in the right direction.
+These two stylised facts suggest that high-income economies use GVCs to outsource low value added downstream production stages and eventually reimport the final goods.
+However, when looking at the development over time it emerges strongly that many developing economies have succeeded in moving up the value chain and that the general trend points to a more even distribution of value added across the different countries.
 
@@
 
+ \appendix
+
+\section{Appendix}
+\begin{table}[h]\small
+  \centering
+   \hspace*{-2.8cm}
+    \begin{tabular}{lrrrrrr}
+    \toprule
+    \multicolumn{1}{l}{\textit{country}} & \multicolumn{1}{c}{\textit{Average (i2e values)}} & \multicolumn{1}{c}{\textit{Average (e2r values)}} & \multicolumn{1}{c}{\textit{Average (i2e)}} & \multicolumn{1}{c}{\textit{Average (e2r)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11 (i2e)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11  (e2r)}} \\
+    \midrule
+    arg   & 52,790 & 66,036 & 12.51\% & 15.65\% & 145.93\% & 30.04\% \\
+    aus   & 178,117 & 343,084 & 13.35\% & 25.71\% & 18.21\% & 59.23\% \\
+    aut   & 250,022 & 214,630 & 25.87\% & 22.21\% & 29.59\% & 39.90\% \\
+    bel   & 437,578 & 285,355 & 32.53\% & 21.22\% & 10.66\% & 30.78\% \\
+    bgr   & 51,393 & 19,864 & 38.01\% & 14.69\% & 32.70\% & 12.35\% \\
+    bra   & 129,301 & 245,839 & 10.95\% & 20.83\% & 37.97\% & 57.77\% \\
+    brn   & 2,412 & 20,438 & 4.51\% & 38.23\% & -41.34\% & 103.27\% \\
+    can   & 647,662 & 407,957 & 23.54\% & 14.83\% & -3.54\% & 70.21\% \\
+    che   & 334,258 & 343,657 & 21.84\% & 22.45\% & 23.45\% & 37.02\% \\
+    chl   & 77,961 & 103,023 & 19.70\% & 26.03\% & 41.95\% & 42.06\% \\
+    chn   & 1,831,434 & 1,293,766 & 24.07\% & 17.00\% & 62.57\% & 37.17\% \\
+    col   & 21,746 & 57,971 & 9.12\% & 24.31\% & -9.63\% & 93.42\% \\
+    cri   & 21,400 & 11,671 & 28.07\% & 15.31\% & 25.42\% & 48.66\% \\
+    cyp   & 12,327 & 8,448 & 22.01\% & 15.09\% & 0.27\% & 53.75\% \\
+    cze   & 290,027 & 129,166 & 41.96\% & 18.69\% & 48.79\% & 11.22\% \\
+    deu   & 1,640,838 & 1,628,409 & 22.51\% & 22.34\% & 71.81\% & 13.06\% \\
+    dnk   & 224,697 & 165,653 & 29.42\% & 21.69\% & 38.07\% & 43.47\% \\
+    esp   & 546,406 & 383,881 & 25.13\% & 17.66\% & 39.96\% & 35.86\% \\
+    est   & 21,777 & 11,992 & 34.90\% & 19.22\% & -3.58\% & 44.23\% \\
+    fin   & 182,478 & 125,397 & 31.43\% & 21.60\% & 44.07\% & 6.99\% \\
+    fra   & 888,006 & 773,925 & 23.01\% & 20.05\% & 44.76\% & 19.68\% \\
+    gbr   & 766,576 & 909,659 & 19.52\% & 23.17\% & 25.71\% & 27.55\% \\
+    grc   & 81,945 & 60,159 & 22.43\% & 16.47\% & 52.51\% & 51.65\% \\
+    hkg   & 115,876 & 121,589 & 18.98\% & 19.92\% & -7.57\% & 53.77\% \\
+    hrv   & 20,725 & 13,277 & 20.09\% & 12.87\% & -3.24\% & -4.93\% \\
+    hun   & 236,208 & 78,585 & 46.20\% & 15.37\% & 59.55\% & 23.83\% \\
+    idn   & 116,161 & 238,302 & 12.97\% & 26.61\% & -4.24\% & 96.09\% \\
+    ind   & 356,692 & 298,471 & 21.34\% & 17.86\% & 178.40\% & 42.28\% \\
+    irl   & 472,729 & 162,263 & 41.96\% & 14.40\% & 11.87\% & 19.69\% \\
+    isl   & 11,301 & 8,977 & 29.32\% & 23.29\% & 84.40\% & 71.61\% \\
+    \end{tabular}
+\end{table}
+
+\begin{table}[h]\small
+  \centering
+    \hspace*{-2.8cm}
+    \begin{tabular}{lrrrrrr}
+    \toprule
+    \multicolumn{1}{l}{\textit{country}} & \multicolumn{1}{c}{\textit{Average (i2e values)}} & \multicolumn{1}{c}{\textit{Average (e2r values)}} & \multicolumn{1}{c}{\textit{Average (i2e)}} & \multicolumn{1}{c}{\textit{Average (e2r)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11 (i2e)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11  (e2r)}} \\
+    \midrule
+    isr   & 105,427 & 75,275 & 23.86\% & 17.04\% & 11.05\% & 53.19\% \\
+    ita   & 778,367 & 641,040 & 23.21\% & 19.12\% & 53.48\% & 35.00\% \\
+    jpn   & 582,907 & 1,388,524 & 11.95\% & 28.47\% & 164.46\% & 32.58\% \\
+    khm   & 11,889 & 3,224 & 37.65\% & 10.21\% & 186.29\% & -36.76\% \\
+    kor   & 1,034,054 & 521,202 & 37.70\% & 19.00\% & 88.43\% & 18.61\% \\
+    ltu   & 16,707 & 15,497 & 22.83\% & 21.17\% & -4.03\% & 42.99\% \\
+    lux   & 237,935 & 51,509 & 53.11\% & 11.50\% & 40.29\% & -15.58\% \\
+    lva   & 14,530 & 13,063 & 25.95\% & 23.33\% & 25.01\% & 34.38\% \\
+    mex   & 479,806 & 214,457 & 28.82\% & 12.88\% & 20.84\% & 31.28\% \\
+    mlt   & 14,762 & 4,931 & 43.26\% & 14.45\% & -27.41\% & 108.31\% \\
+    mys   & 517,084 & 215,738 & 41.45\% & 17.29\% & 33.36\% & 23.93\% \\
+    nld   & 306,010 & 390,300 & 19.50\% & 24.87\% & -14.29\% & 51.12\% \\
+    nor   & 163,813 & 351,592 & 16.78\% & 36.01\% & -13.46\% & 57.48\% \\
+    nzl   & 39,712 & 33,912 & 17.08\% & 14.59\% & -0.73\% & 48.61\% \\
+    phl   & 103,838 & 82,783 & 29.04\% & 23.15\% & -20.59\% & 101.26\% \\
+    pol   & 273,027 & 198,877 & 29.34\% & 21.37\% & 99.99\% & 15.65\% \\
+    prt   & 125,283 & 64,391 & 30.77\% & 15.81\% & 18.69\% & 37.51\% \\
+    rou   & 59,385 & 54,956 & 24.23\% & 22.42\% & 14.95\% & 45.54\% \\
+    rus   & 317,701 & 837,747 & 13.51\% & 35.62\% & 3.61\% & 56.47\% \\
+    sau   & 57,392 & 547,987 & 3.95\% & 37.73\% & -15.39\% & 56.02\% \\
+    sgp   & 548,286 & 219,149 & 46.08\% & 18.42\% & 12.95\% & 61.94\% \\
+    svk   & 140,548 & 61,023 & 44.95\% & 19.52\% & 47.60\% & 7.80\% \\
+    svn   & 51,465 & 29,065 & 34.78\% & 19.64\% & 11.26\% & 58.26\% \\
+    swe   & 355,353 & 262,980 & 29.09\% & 21.52\% & 8.68\% & 29.44\% \\
+    tha   & 391,773 & 156,527 & 36.05\% & 14.40\% & 61.19\% & 23.84\% \\
+    tun   & 35,124 & 18,724 & 30.17\% & 16.08\% & 30.65\% & 48.33\% \\
+    tur   & 180,927 & 113,136 & 22.30\% & 13.95\% & 195.25\% & 12.71\% \\
+    twn   & 649,797 & 353,241 & 39.52\% & 21.48\% & 41.83\% & 60.53\% \\
+    usa   & 1,318,846 & 2,248,028 & 13.52\% & 23.04\% & 30.75\% & 26.87\% \\
+    vnm   & 119,821 & 57,005 & 33.76\% & 16.06\% & 72.69\% & 19.95\% \\
+    zaf   & 102,394 & 122,842 & 19.31\% & 23.16\% & 47.60\% & 24.98\% \\
+    \bottomrule
+    \end{tabular}
+      \caption{GVC indicators by country}
+  \label{tab:gvc_k}
+\end{table}
+
+\begin{table}[h]\small
+  \centering
+    \hspace*{-3cm}
+    \begin{tabular}{lrrrrrr}
+    \toprule
+    \multicolumn{1}{l}{\textit{sector}} & \multicolumn{1}{c}{\textit{Average (i2e values)}} & \multicolumn{1}{c}{\textit{Average (e2r values)}} & \multicolumn{1}{c}{\textit{Average (i2e)}} & \multicolumn{1}{c}{\textit{Average (e2r)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11 (i2e)}} & \multicolumn{1}{c}{\textit{$\Delta$ 95-11  (e2r)}} \\
+    \midrule
+    c01t05agr & 227,969 & 364,681 & 13.00\% & 20.79\% & 36.13\% & 27.51\% \\
+    c10t14min & 435,816 & 3,324,446 & 5.82\% & 44.38\% & -4.51\% & 55.66\% \\
+    c15t16fod & 675,902 & 146,824 & 19.58\% & 4.25\% & 23.04\% & 32.73\% \\
+    c17t19tex & 679,185 & 174,555 & 23.30\% & 5.99\% & 9.49\% & 2.72\% \\
+    c20wod & 105,771 & 80,461 & 20.56\% & 15.64\% & 30.59\% & 63.12\% \\
+    c21t22pap & 302,255 & 344,112 & 19.23\% & 21.89\% & 27.63\% & 7.10\% \\
+    c23pet & 1,285,522 & 356,703 & 39.53\% & 10.97\% & 65.81\% & -8.04\% \\
+    c24chm & 1,762,631 & 925,255 & 28.18\% & 14.79\% & 52.19\% & -2.18\% \\
+    c25rbp & 446,528 & 309,253 & 27.89\% & 19.32\% & 38.26\% & 3.06\% \\
+    c26nmm & 161,574 & 126,800 & 22.09\% & 17.34\% & 41.44\% & 17.72\% \\
+    c27met & 1,340,507 & 868,120 & 31.03\% & 20.10\% & 36.53\% & -11.26\% \\
+    c28fbm & 481,732 & 441,396 & 27.53\% & 25.23\% & 38.11\% & 6.86\% \\
+    c29meq & 1,398,864 & 570,006 & 26.50\% & 10.80\% & 39.73\% & 30.96\% \\
+    c303233ceq & 3,162,705 & 1,062,750 & 38.63\% & 12.98\% & 45.28\% & 18.74\% \\
+    c31elq & 713,928 & 301,827 & 30.71\% & 12.98\% & 37.32\% & -3.42\% \\
+    c34mtr & 1,827,519 & 340,198 & 34.53\% & 6.43\% & 33.39\% & 7.25\% \\
+    c35trq & 745,063 & 203,834 & 30.54\% & 8.35\% & 40.89\% & 10.73\% \\
+    c36t37otm & 429,561 & 156,828 & 23.49\% & 8.58\% & 14.69\% & 66.57\% \\
+    c50t52wrt & 923,756 & 3,002,076 & 9.15\% & 29.75\% & 35.67\% & 33.76\% \\
+    c60t63trn & 1,293,729 & 1,507,876 & 17.88\% & 20.84\% & 60.51\% & 33.82\% \\
+    c64ptl & 81,529 & 322,823 & 12.39\% & 49.07\% & 85.71\% & -7.06\% \\
+    c65t67fin & 371,026 & 992,980 & 13.51\% & 36.17\% & 85.73\% & -8.65\% \\
+    c71rmq & 67,266 & 215,714 & 11.59\% & 37.18\% & 89.87\% & 13.54\% \\
+    c72its & 144,310 & 286,414 & 16.55\% & 32.86\% & 73.87\% & -1.85\% \\
+    c73t74bzs & 386,997 & 1,922,031 & 11.27\% & 55.99\% & 44.41\% & 17.81\% \\
+    \bottomrule
+    \end{tabular}
+      \caption{GVC indicators by sector}
+  \label{tab:gvc_s}
+\end{table}
+
 \end{document}
